---
title: "AllSierras_Historical"
output: html_document
---

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(dataRetrieval)
library(sf)
library(maps)
library(sp)
library(ncdf4)
library(sf)
library(sp)
library(lwgeom)
library(terra)
library(raster)
library(mapview)
library(tidyterra)
library(viridis)
library(MetBrewer)
library(stars)
library(data.table)
library(tigris)
library(patchwork)
library(car)
library(multcomp)
```

#Input Files (all together)
```{r}
AllSierras_DamCapacity <- readRDS("AllSierras_DamCapacity_volume.rds")
PADUS <- readRDS("PADUS.rds")
WatershedBoundaries <- readRDS("WatershedBoundaries.rds")
PADUS_clip <- readRDS("PADUS_clip.rds")
PADUS_Gap12 <- readRDS("PADUS_Gap12.rds")
DamCap_GAP12 <- readRDS("DamCap_GAP12.rds")
TNC_DamCapacity <- readRDS("TNC_DamCapacity.rds")
fireRisk <- readRDS("fireRisk.rds")
tribalLands <- readRDS("tribalLands.rds")
contFireRisk <- readRDS("contFireRisk.rds")


```

##Watershed Boundaries Map
```{r}
palette_new <- colorRampPalette(colors = c("#ff0000", "#ffa500", "#ffff00", "#00FF50", "#4900FF"))(5)

w <- ggplot() +
    #geom_sf(data = background, fill = "#dcdcdc", lwd = 0.1)+
    geom_sf(data = WatershedBoundaries, fill = "#b0b0b0", lwd = 0.1, alpha = 0.5 ) +
    scale_color_manual(values=palette_new)+
    theme_minimal()+
    ggtitle(expression(paste("Watershed Boundaries"))) +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    #theme(element_text(size=10),legend.position = c(.15,.2))+
    #labs(color = "dams/km", size=10)+
    theme(panel.border = element_rect(colour = "black", fill=NA))

```
##Current Dam Capacity Map
```{r}
#scales::show_col(palette_new)

c <- ggplot() +
    #geom_sf(data = background, fill = "#dcdcdc", lwd = 0.1)+
    geom_sf(data = transform(AllSierras_DamCapacity, DamCapacity=cut(oCC_EX, breaks = c(0,.1,2,5,15,40), include.lowest=T)),aes(color = DamCapacity)) +
    scale_color_manual(values=palette_new)+
    theme_minimal()+
    ggtitle(expression(paste("Current Dam Capacity"))) +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    theme(element_text(size=10),legend.position = c(.15,.2))+
    labs(color = "dams/km", size=10)+
    theme(panel.border = element_rect(colour = "black", fill=NA))

```
##Historical Dam Capacity Map
```{r}
h <- ggplot() +
    #geom_sf(data = background, fill = "#dcdcdc", lwd = 0.1)+
    geom_sf(data = transform(AllSierras_DamCapacity, DamCapacity=cut(oCC_HPE, breaks = c(0,.1,2,5,15,40), include.lowest=T)),aes(color = DamCapacity)) +
    scale_color_manual(values=palette_new)+
    theme_minimal()+
    ggtitle(expression(paste("Historical Dam Capacity"))) +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    theme(element_text(size=10),legend.position = c(.15,.2))+
    labs(color = "dams/km", size=10)+
    theme(panel.border = element_rect(colour = "black", fill=NA))

```
##Both Current and Historical Dam Capacity Maps
```{r}
(h + c)
```
#Calculate Historical Percent Remaining
```{r}
AllSierras_DamCapacity$mCC_EXvHPE <- AllSierras_DamCapacity$oCC_EX/AllSierras_DamCapacity$oCC_HPE
#When doing this division, sometimes we get 0/0 = Inf, especially in areas that are slope limited. In one way of looking at it, these values should be converted to 1, since dam capacity stayed the same from historic to current. However we converted them to NA? 
AllSierras_DamCapacity$mCC_EXvHPE[!is.finite(AllSierras_DamCapacity$mCC_EXvHPE)] <- NA
summary(AllSierras_DamCapacity$mCC_EXvHPE)
```

##Summarize Historical Percent Remaining
```{r}
mean(AllSierras_DamCapacity$mCC_EXvHPE, na.rm=T)#60% dam cap remaining by capacity 
existingsum <- sum(AllSierras_DamCapacity$mCC_EX_CT)
historicalsum <- sum(AllSierras_DamCapacity$mCC_HPE_CT)
existingsum/historicalsum # 51% dam cap remaining by count

```

##Percent Remaining Map
```{r}
r <- ggplot() +
    #geom_sf(data = background, fill = "#dcdcdc", lwd = 0.1)+
    geom_sf(data = transform(AllSierras_DamCapacity, DamCapacity=cut(mCC_EXvHPE, breaks = c(0,.25,.5,.75,1,70), include.lowest=T)),aes(color = DamCapacity)) +
    scale_color_manual(values=palette_new)+
    theme_minimal()+
    ggtitle(expression(paste("Percent Dam Capacity Remaining"))) +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    theme(element_text(size=10),legend.position = c(.15,.2))+
    labs(color = "dams/km", size=10)+
    theme(panel.border = element_rect(colour = "black", fill=NA))
r
```
##Ttest Percent Dam Cap remaining, GAP and Other
```{r}
AllSierras_DamCapacity <- st_join(AllSierras_DamCapacity, PADUS_Gap12)

```

```{r}

AllSierras_DamCapacity$Protected <- with(AllSierras_DamCapacity, ifelse(is.na(AllSierras_DamCapacity$GAP_Sts), "Other", "GAP 1 & 2"))
AllSierras_DamCapacity$Protected <- as.factor(AllSierras_DamCapacity$Protected)
       
leveneTest(mCC_EXvHPE~Protected, data = AllSierras_DamCapacity)#significant
ttest_1 <- t.test(mCC_EXvHPE ~ Protected, data = AllSierras_DamCapacity)
ttest_1

```

##Total dams lost
```{r}
gap <- subset(AllSierras_DamCapacity, Protected == "GAP 1 & 2")
other <- subset(AllSierras_DamCapacity, Protected == "Other")

p1 <- ggplot()+
  geom_boxplot(data = gap, aes(x=Protected, y = sqrt(mCC_EXvHPE)), fill = "#e78429")+
  geom_boxplot(data = other, aes(x=Protected, y = sqrt(mCC_EXvHPE)), fill= "grey60")+
  theme_minimal() +
  ggtitle(expression(paste("% Change in Dam Capacity"))) + 
  xlab("Protection Status")+
  ylab("Sqrt % Change in Dam Capacity")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))

```

##ANOVA within GAP 12
```{r}
###Clip BRAT to GAP 1 and 2 area
DamCap_GAP12 = st_intersection(AllSierras_DamCapacity, PADUS_Gap12)
DamCap_GAP12$FeatClass <- as.factor(DamCap_GAP12$FeatClass)
                                    
ggplot(data = DamCap_GAP12, aes(x=FeatClass, y = log(mCC_EXvHPE)))+
         geom_boxplot()+
         theme_minimal()
       
leveneTest(mCC_EXvHPE~FeatClass, data = DamCap_GAP12)#significant
anova1 <- aov(mCC_EXvHPE ~ FeatClass, data = DamCap_GAP12)
anova1_residuals <- residuals(object = anova1)

ggplot(anova1, aes(x = .fitted, y = .resid))+
  geom_point() + theme_classic()#bad
#shapiro.test(x = anova1_residuals)#too big

#try square root transformation
DamCap_GAP12 <- DamCap_GAP12 %>% mutate(sqrtvalue=sqrt(mCC_EXvHPE))
anova2 <- aov(sqrtvalue ~ FeatClass, data = DamCap_GAP12)

ggplot(data = DamCap_GAP12, aes(x=FeatClass, y = (sqrtvalue)))+
         geom_boxplot()+
         theme_minimal()

leveneTest(sqrtvalue~FeatClass, data = DamCap_GAP12)#significant

kruskal.test(sqrtvalue ~ FeatClass, data = DamCap_GAP12)

#PostHoc on the square root transformation
summary(anova2)
tukey <- TukeyHSD(anova2)
print(tukey)
```

```{r}
desig <- subset(DamCap_GAP12, FeatClass == "Designation")
fee <- subset(DamCap_GAP12, FeatClass == "Fee")
easement <- subset(DamCap_GAP12, FeatClass == "Easement")


p2 <- ggplot()+
  geom_boxplot(data = desig, aes(x=FeatClass, y = sqrt(mCC_EXvHPE)), fill = "#88a0dc") +
  geom_boxplot(data = fee, aes(x=FeatClass, y = sqrt(mCC_EXvHPE)), fill = "#ed968c") +
  geom_boxplot(data = easement, aes(x=FeatClass, y = sqrt(mCC_EXvHPE)), fill = "#f9d14a") +
  theme_minimal() +
  ggtitle(expression(paste("% Change in Dam Capacity"))) + 
  xlab("Land Ownership Type")+
  ylab("Sqrt % Change in Dam Capacity")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))

```
##Both boxplots together
```{r}
p1 + p2
```
#Volume ~ Dam Capacity
```{r}
lm_volume <- lm(AllSierras_DamCapacity$volume ~ AllSierras_DamCapacity$oCC_EX)
lm_volume
```

```{r}
ggplot()+
  geom_point(data = AllSierras_DamCapacity, aes(x = oCC_EX, y = volume))+
  #stat_summary(fun.data=lm_volume) + 
  geom_smooth(method='lm', formula= x~y) +
  theme_minimal() +
  ggtitle(expression(paste("Volume ~ Dam Capacity"))) + 
  xlab("Current Dam Capacity (dams/km)")+
  ylab("Potential Water Storage (m3)")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  #labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))
hist(AllSierras_DamCapacity$oCC_EX, breaks = 160)
```


#Change in veg suitability
```{r}
AllSierras_DamCapacity$VegChange100 <-  AllSierras_DamCapacity$iVeg100EX - AllSierras_DamCapacity$iVeg100Hpe 

```

##Change in veg suitability map
```{r}

veg <- ggplot() +
    geom_sf(data = AllSierras_DamCapacity, aes(color = VegChange100)) +
    scale_color_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0, na.value = "grey70")+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Vegetation Change"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) 
```
#Change in volume
##Historical volume calculation
```{r}
for (i in AllSierras_DamCapacity){ #for each stream segment in the BRAT model output shapefile
  damCt <- AllSierras_DamCapacity$mCC_HPE_CT #number of beaver dams per segment (calculated based on dam density and segment length as the column mCC_HPE_CT in the attribute table)
  slope <- AllSierras_DamCapacity$iGeo_Slope #meters
  da <- AllSierras_DamCapacity$iGeo_DA #m2
  width <- ifelse(da < 4.95,2.18*(da)^0.191,
                  ifelse(da >= 4.95 & da < 377,1.41*(da)^0.462,7.18*(da)^0.183)) #based on Scamardo et al. 2022
  height <- 1 #assume all beaver dams are 1m tall
  length <- height/slope #trigonometry: slope = rise/run, so run = rise/slope
  maxLength<- (AllSierras_DamCapacity$iGeo_Len/damCt) #set a maximum length so that we can change volume calculations if there are so many dams that they all run together
  volumeHist <- ifelse(length < maxLength, 0.5*(height)*(width)*(length)*(damCt),(height)*(width)*(maxLength)*(damCt))
  surfaceAreaHist <- ifelse(length < maxLength, (width)*(length)*(damCt),(width)*(maxLength)*(damCt))
  AllSierras_DamCapacity$width <- width
  AllSierras_DamCapacity$volumeHist <- volumeHist #m3
  AllSierras_DamCapacity$surfaceAreaHist <- surfaceAreaHist #m2
} 

```

##Historical volume map
```{r}
ggplot() +
    geom_sf(data = AllSierras_DamCapacity, aes(color = volumeHist)) +
    scale_color_viridis(discrete = FALSE, option = "plasma", na.value = "grey70")+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Potential Water Storage (m3)"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) 
```
##Volume Change
```{r}
AllSierras_DamCapacity$VolumeChange <-  AllSierras_DamCapacity$volume - AllSierras_DamCapacity$volumeHist

```
##Volume change map
```{r}
vol <- ggplot() +
    geom_sf(data = AllSierras_DamCapacity, aes(color = VolumeChange)) +
    scale_color_gradient2(low = "red", mid = "grey70", high = "blue", midpoint = 0, na.value = "grey70")+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Change in Volume (m3)"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) 
```

##Volume change + Veg change in the same frame
```{r}
vol + veg
```
##Summarize with numbers
```{r}
sum(AllSierras_DamCapacity$VolumeChange )
currentstor <- sum(AllSierras_DamCapacity$volume)
historicalstor <- sum(AllSierras_DamCapacity$volumeHist)
percentleft <- currentstor/historicalstor

volchange <- currentstor-historicalstor
```



#Apply volume to every watershed
```{r}
watersheds <- st_join(WatershedBoundaries,AllSierras_DamCapacity)
watershedsPoly <- st_intersection(AllSierras_DamCapacity, WatershedBoundaries)
saveRDS(watershedsPoly,"watershedsPoly.rds")
saveRDS(watersheds,"watersheds.rds")

test <- watersheds %>%
    group_by(geometry) %>% 
    summarise(totalVol = sum(volume),
              totalVolHist = sum(volumeHist),
              totalVolChange = sum(VolumeChange))

totalVol <- ggplot() +
  geom_sf(data = WatershedBoundaries) +
  geom_sf(data = test, aes(fill = totalVol), inherit.aes = FALSE) +
  scale_fill_viridis_c(option = "plasma") + 
  xlab(NULL)+
  ylab(NULL)+
  ggtitle(expression(paste("Potential Water Storage by Watershed (m3)"))) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
  theme(panel.border = element_rect(colour = "black", fill=NA))


```
#Apply Dam Capacity to every watershed
```{r}

test <- watersheds %>%
    group_by(geometry) %>% 
    summarise(totalVol = sum(volume),
              totalVolHist = sum(volumeHist),
              totalVolChange = sum(VolumeChange),
              totalBeavs = sum(mCC_EX_CT ),
              avgBeavs = mean(oCC_EX))

totalBeavs <- ggplot() +
  geom_sf(data = WatershedBoundaries) +
  geom_sf(data = test, aes(fill = totalBeavs), inherit.aes = FALSE) +
  scale_fill_viridis_c() + 
  xlab(NULL)+
  ylab(NULL)+
  ggtitle(expression(paste("Total dam building capacity (# of dams)"))) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5))+
  theme(panel.border = element_rect(colour = "black", fill=NA))


```

```{r}
totalVol + totalBeavs
```


#Limitations
```{r}
unique(AllSierras_DamCapacity$oPBRC_UD)
length(AllSierras_DamCapacity$oPBRC_UD)


DamBuildingPossible <- (length(AllSierras_DamCapacity$oPBRC_UD[AllSierras_DamCapacity$oPBRC_UD == "Dam Building Possible"]))/(length(AllSierras_DamCapacity$oPBRC_UD))

StreamPowerLimited <- (length(AllSierras_DamCapacity$oPBRC_UD[AllSierras_DamCapacity$oPBRC_UD == "Stream Power Limited"]))/(length(AllSierras_DamCapacity$oPBRC_UD))

NaturallyVegetationLimited <- (length(AllSierras_DamCapacity$oPBRC_UD[AllSierras_DamCapacity$oPBRC_UD == "Naturally Vegetation Limited"]))/(length(AllSierras_DamCapacity$oPBRC_UD))

StreamSizeLimited <- (length(AllSierras_DamCapacity$oPBRC_UD[AllSierras_DamCapacity$oPBRC_UD == "Stream Size Limited"]))/(length(AllSierras_DamCapacity$oPBRC_UD))

PotentialReservoir <- (length(AllSierras_DamCapacity$oPBRC_UD[AllSierras_DamCapacity$oPBRC_UD == "Potential Reservoir or Landuse"]))/(length(AllSierras_DamCapacity$oPBRC_UD))

SlopeLimited <- (length(AllSierras_DamCapacity$oPBRC_UD[AllSierras_DamCapacity$oPBRC_UD == "Slope Limited"]))/(length(AllSierras_DamCapacity$oPBRC_UD))

AnthropogenicallyLimited <- (length(AllSierras_DamCapacity$oPBRC_UD[AllSierras_DamCapacity$oPBRC_UD == "Anthropogenically Limited"]))/(length(AllSierras_DamCapacity$oPBRC_UD))

a <- c('DamBuildingPossible','StreamPowerLimited','NaturallyVegetationLimited','StreamSizeLimited','PotentialReservoir','SlopeLimited','AnthropogenicallyLimited') 
b <- c(DamBuildingPossible,StreamPowerLimited,NaturallyVegetationLimited,StreamSizeLimited,PotentialReservoir,SlopeLimited,AnthropogenicallyLimited) 
df <- data.frame(a,b)

colnames(df) <- c("Limitation", "Proportion")
df$Percent <- df$Proportion * 100
df
```
##Map
```{r}
palette_discrete <- colorRampPalette(colors = c("#E41A1C", "#CCCCCC", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#0070FF"))(7)
#scales::show_col(palette_new)

ggplot() +
    geom_sf(data = AllSierras_DamCapacity,aes(color = oPBRC_UD)) +
    scale_color_manual(values=palette_discrete)+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Restoration Limitations"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) 

```


#Fire Risk
##Clip
```{r}
fireRisk = st_intersection(WatershedBoundaries, fireRisk)
```
##Join Fire Risk with Dam Capacity
```{r}
AllSierras_DamCapacity <- st_join(AllSierras_DamCapacity, fireRisk)
```

##Pull out individual designations
```{r}
unique(AllSierras_DamCapacity$HAZ_CLASS)
moderate <- AllSierras_DamCapacity[AllSierras_DamCapacity$HAZ_CLASS=="Moderate",]
high <- AllSierras_DamCapacity[AllSierras_DamCapacity$HAZ_CLASS=="High",]
veryHigh <- AllSierras_DamCapacity[fireRisk$HAZ_CLASS=="Very High",]

```

##Map
```{r}
fire <- ggplot() + 
  #geom_sf(data = background, fill = "#dcdcdc", lwd = 0.1)+
  geom_sf(data = moderate, fill = "#ffff00", color = NA, aes(shape = "Moderate")) +
  geom_sf(data = high, fill = "#ffa500",color = NA, alpha = 0.8, aes(shape = "High")) +
  geom_sf(data = veryHigh, fill = "#ff0000", color = NA, aes(shape = "Very High")) +
  coord_sf(crs = st_crs(3310))+ 
  theme_minimal() +
  ggtitle(expression(paste("Fire Risk (State Responsibility Areas)"))) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  theme(legend.position = c(.25,.15))+
  labs(shape = "")+
  #coord_sf(xlim = c(-124.0, -116.8), ylim = c(34, 41)) +
  theme(panel.border = element_rect(colour = "black", fill=NA))



```

##ANOVA fire risk Dam Capacity
```{r}
AllSierras_DamCapacity$HAZ_CLASS <- as.factor(AllSierras_DamCapacity$HAZ_CLASS)
                                    
ggplot(data = AllSierras_DamCapacity, aes(x=HAZ_CLASS, y = log(oCC_EX)))+
         geom_boxplot()+
         theme_minimal()
       
leveneTest(oCC_EX~HAZ_CLASS, data = AllSierras_DamCapacity)#significant
anova_fire <- aov(oCC_EX ~ HAZ_CLASS, data = AllSierras_DamCapacity)
anova_fire_resid <- residuals(object = anova_fire)

ggplot(anova_fire_resid, aes(x = .fitted, y = .resid))+
  geom_point() + theme_classic()#fine
#shapiro.test(x = anova1_residuals)#too big

kruskal.test(oCC_EX ~ HAZ_CLASS, data = AllSierras_DamCapacity)

#PostHoc on the square root transformation
summary(anova_fire)
tukey <- TukeyHSD(anova_fire)
print(tukey)
```

```{r}
moderate <- subset(AllSierras_DamCapacity, HAZ_CLASS == "Moderate")
high <- subset(AllSierras_DamCapacity, HAZ_CLASS == "High")
veryHigh <- subset(AllSierras_DamCapacity, HAZ_CLASS == "Very High")

  
f1 <- ggplot()+
  geom_boxplot(data = moderate, aes(x=HAZ_CLASS, y = oCC_EX), fill = "#ffff00") +
  geom_boxplot(data = high, aes(x=HAZ_CLASS, y = oCC_EX), fill = "#ffa500") +
  geom_boxplot(data = veryHigh, aes(x=HAZ_CLASS, y = oCC_EX), fill = "#ff0000") +
  theme_minimal() +
  ggtitle(expression(paste("Dam Capacity by Fire Risk"))) + 
  xlab("Fire Risk")+
  ylab("Dam Capacity (dams/km)")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))

```
##ANOVA fire risk % change Dam Capacity
```{r}
AllSierras_DamCapacity$HAZ_CLASS <- as.factor(AllSierras_DamCapacity$HAZ_CLASS)
                                    
ggplot(data = AllSierras_DamCapacity, aes(x=HAZ_CLASS, y = log(mCC_EXvHPE)))+
         geom_boxplot()+
         theme_minimal()
       
leveneTest(mCC_EXvHPE~HAZ_CLASS, data = AllSierras_DamCapacity)#significant
anova_fire <- aov(mCC_EXvHPE ~ HAZ_CLASS, data = AllSierras_DamCapacity)
anova_fire_resid <- residuals(object = anova_fire)

ggplot(anova_fire_resid, aes(x = .fitted, y = .resid))+
  geom_point() + theme_classic()#fine
#shapiro.test(x = anova1_residuals)#too big

kruskal.test(mCC_EXvHPE ~ HAZ_CLASS, data = AllSierras_DamCapacity)

#PostHoc on the square root transformation
summary(anova_fire)
tukey <- TukeyHSD(anova_fire)
print(tukey)
```

```{r}
f2 <- ggplot()+
  geom_boxplot(data = moderate, aes(x=HAZ_CLASS, y = sqrt(mCC_EXvHPE)), fill = "#ffff00") +
  geom_boxplot(data = high, aes(x=HAZ_CLASS, y = sqrt(mCC_EXvHPE)), fill = "#ffa500") +
  geom_boxplot(data = veryHigh, aes(x=HAZ_CLASS, y = sqrt(mCC_EXvHPE)), fill = "#ff0000") +
  theme_minimal() +
  ggtitle(expression(paste("% Change in Dam Capacity by Fire Risk"))) + 
  xlab("Fire Risk")+
  ylab("Sqrt % Change in Dam Capacity")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))

```
```{r}
f1 + f2
```
#Fire Risk Continuous
```{r}
# Read in the raster (already projected and clipped in ArcPro)
contFireRisk <- rast("C:/Users/jessie_moravek/Desktop/SierraNevadaBeavers/FireRisk/FireRisk_Clip.tif")
plot(contFireRisk)

#visualie
contFireRisk_df <- as.data.frame(contFireRisk, xy=T)
ggplot() +
    geom_raster(data = contFireRisk_df , aes(x = x, y = y, fill = Band_1)) +
    scale_fill_viridis_c(option = "inferno") +
    coord_quickmap()
hist(contFireRisk_df$Band_1)
saveRDS(contFireRisk,"contFireRisk.rds")

#The WHP has continous "fire risk values" that in this area go from 0 to 42223. Those are an index developed to indiccate Wildfire Hazard Potential (WHP). See metadata for more information. 
#This is a continuous dataset but much like BRAT is often presented as categorical for visualization purposes. 
```



#Large Complexes
```{r}
complex <- AllSierras_DamCapacity[AllSierras_DamCapacity$mCC_EX_CT >5,]
complexHist <- AllSierras_DamCapacity[AllSierras_DamCapacity$mCC_HPE_CT >5,]
```

```{r}
nrow(complex) / nrow(AllSierras_DamCapacity)
nrow(complexHist) / nrow(AllSierras_DamCapacity)
(sum(complex$Shape_Leng.x))*.001
(sum(complexHist$Shape_Leng.x))*.001

```


####Map
```{r}
palette_new <- colorRampPalette(colors = c( "#00FF50", "#4900FF"))(2)
#scales::show_col(palette_new)

ggplot() +
    geom_sf(data = complex, color = "#4900FF") +
    scale_color_manual(values=palette_new)+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Complexes >5 dams"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(colour = "black", fill=NA))

```



#Tribal Lands
##Clip
```{r}
tribalLands = st_intersection(WatershedBoundaries, tribalLands)
```
##Join Tribal Lands with Dam Capacity
```{r}
AllSierras_DamCapacity <- st_join(AllSierras_DamCapacity, tribalLands)
```


##Map
```{r}
tl <- ggplot() + 
  #geom_sf(data = background, fill = "#dcdcdc", lwd = 0.1)+
  geom_sf(data = WatershedBoundaries, fill = "#dcdcdc",alpha = 0.5 ) +
  geom_sf(data = tribalLands, fill = "#ff0000", lwd = 0.1) +
  coord_sf(crs = st_crs(3310))+ 
  theme_minimal() +
  ggtitle(expression(paste("Federally Recognized Tribal Lands"))) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  theme(legend.position = c(.25,.15))+
  labs(shape = "")+
  #coord_sf(xlim = c(-124.0, -116.8), ylim = c(34, 41)) +
  theme(panel.border = element_rect(colour = "black", fill=NA))

```

#Restoration opporunities
```{r}
DamCap_RestOpp <- AllSierras_DamCapacity[AllSierras_DamCapacity$oPBRC_CR != "NA",]

```

```{r}
palette_discrete <- colorRampPalette(colors = c("#E41A1C", "#CCCCCC", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#0070FF"))(7)
#scales::show_col(palette_new)

ggplot() +
    geom_sf(data = AllSierras_DamCapacity,aes(color = oPBRC_CR)) +
    scale_color_manual(values=palette_discrete)+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Restoration Opportunities"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) 

```



