---
title: "AllSierras_DamHeights"
output: html_document
---
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(dataRetrieval)
library(sf)
library(maps)
library(sp)
library(ncdf4)
library(sf)
library(sp)
library(lwgeom)
library(terra)
library(raster)
library(mapview)
library(tidyterra)
library(viridis)
library(MetBrewer)
library(stars)
library(data.table)
```
#ALL STREAMS
##Prep Work
###NASA projection
```{r}
wkt <- 'GEOGCS["WGS 84",DATUM["WGS_1984",SPHEROID["WGS 84",6378137,298.257223563,AUTHORITY["EPSG","7030"]],AUTHORITY["EPSG","6326"]],PRIMEM["Greenwich",0,AUTHORITY["EPSG","8901"]],UNIT["degree",0.0174532925199433,AUTHORITY["EPSG","9122"]],AXIS["Latitude",NORTH],AXIS["Longitude",EAST],AUTHORITY["EPSG","4326"]]'
```


###Dam Capacity
```{r}
# Read in the shapefile with st_read()
AllSierras_DamCapacity <- st_read("C:/Users/jessie_moravek/Desktop/SierraNevadaBeavers/AllSierras_DamCapacity2.shp")

#remove unneeded shapefiles
file.remove(list.files(pattern = "AllSierras_DamCapacity*",recursive = F))

# Read in shapefile from previously imported shapefile
AllSierras_DamCapacity <- readRDS("AllSierras_DamCapacity.rds")

#Project to wkt (NASA projection)
AllSierras_DamCapacity <- st_transform(AllSierras_DamCapacity, 3310)

AllSierras_DamCapacity <- st_transform(AllSierras_DamCapacity, wkt)
saveRDS(AllSierras_DamCapacity,"AllSierras_DamCapacity.rds")
```
####Map
```{r}
palette_new <- colorRampPalette(colors = c("#ff0000", "#ffa500", "#ffff00", "#00FF50", "#4900FF"))(5)
#scales::show_col(palette_new)

ggplot(transform(AllSierras_DamCapacity, DamCapacity=cut(oCC_EX, breaks = c(0,.1,2,5,15,40), include.lowest=T))) +
    geom_sf(aes(color = DamCapacity)) +
    scale_color_manual(values=palette_new)+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Dam Capacity (Dams/km)"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) 

```
####Map
```{r}
palette_discrete <- colorRampPalette(colors = c("#E41A1C", "#CCCCCC", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#0070FF"))(7)
#scales::show_col(palette_new)

ggplot() +
    geom_sf(data = AllSierras_DamCapacity,aes(color = oPBRC_UD)) +
    scale_color_manual(values=palette_discrete)+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Restoration Limitations"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) 

```

###Make a grid
#Note: Can't make a bounding box in a non-metric projection without first running this through LIS so we know what the LIS model wants the dimensions to be so we can set them for the grid. Need to have Augusto make an input file for us before we can proceed. 
```{r}
#Make a bounding box extent
#longitude	latitude
#-121.605	34.34
#-117.305	40.495
xmin <- -121.605
xmax <- -117.305
ymin <- 34.340
ymax <- 40.495
extent <- extent(xmin, xmax, ymin, ymax)

#Make a bounding box in the correct projection
study_area <- st_bbox(extent, crs = st_crs(wkt))

#Is wkt in metric or lat/long coordinate system? st_is_longlat gives TRUE, so it is NOT in metric.
st_is_longlat(study_area)

#Make a grid of the correct size
study_area = st_make_grid(study_area, n = c(174,63))

#Convert to an sf dataframe for future manipulation
study_area = st_as_sf(study_area)

#Add an ID column
study_area$id = 1:nrow(study_area)
```

###Volume calculation
```{r}
for (i in AllSierras_DamCapacity){ #for each stream segment in the BRAT model output shapefile
  damCt <- AllSierras_DamCapacity$mCC_EX_CT #number of beaver dams per segment (calculated based on dam density and segment length as the column mCC_EX_CT in the attribute table)
  slope <- AllSierras_DamCapacity$iGeo_Slope #meters
  da <- AllSierras_DamCapacity$iGeo_DA #m2
  width <- ifelse(da < 4.95,2.18*(da)^0.191,
                  ifelse(da >= 4.95 & da < 377,1.41*(da)^0.462,7.18*(da)^0.183)) #based on Scamardo et al. 2022
  height <- 1 #assume all beaver dams are 1m tall
  length <- height/slope #trigonometry: slope = rise/run, so run = rise/slope
  maxLength<- (AllSierras_DamCapacity$iGeo_Len/damCt) #set a maximum length so that we can change volume calculations if there are so many dams that they all run together
  volume <- ifelse(length < maxLength, 0.5*(height)*(width)*(length)*(damCt),(height)*(width)*(maxLength)*(damCt))
  surfaceArea <- ifelse(length < maxLength, (width)*(length)*(damCt),(width)*(maxLength)*(damCt))
  AllSierras_DamCapacity$width <- width
  AllSierras_DamCapacity$volume <- volume #m3
  AllSierras_DamCapacity$surfaceArea <- surfaceArea #m2
} 


saveRDS(AllSierras_DamCapacity,"AllSierras_DamCapacity_volume.rds")
```

###Volume by grid cell
```{r}
sf_use_s2(FALSE)
intersection <- st_intersection(study_area, damCapacity2_v) %>% 
  mutate(length = st_length(.)) %>% #creates a "length" column that is the length of that particular segment in the grid cell from the "id" column
  st_drop_geometry() # complicates things in joins later on

study_area <- study_area %>% 
  left_join(intersection, by = "id")

vol_by_grid <- study_area %>% 
  mutate(percent_len_in_cell = length/iGeo_Len) %>% 
  mutate(segment_vol_in_cell = percent_len_in_cell*volume) %>% 
  group_by(id) %>% summarize(total_vol_in_cell = sum(segment_vol_in_cell))

vol_by_grid$total_vol_in_cell <- as.numeric(vol_by_grid$total_vol_in_cell) #convert volume col to numeric

saveRDS(vol_by_grid,"vol_by_grid.rds")
vol_by_grid <- readRDS("vol_by_grid.rds")

```
###Summary
```{r}
summary(vol_by_grid)
sum(vol_by_grid$total_vol_in_cell, na.rm=T)
```

####Map
```{r}
ggplot() +
    geom_sf(data = vol_by_grid, aes(fill = total_vol_in_cell), color = NA) +
    #geom_sf(data = damCapacity2, col = "blue") +
    scale_fill_gradient2(low = "gray70", high = "#078FF7") +
    xlab(NULL)+
    ylab(NULL)+
    theme_minimal()+
    ggtitle(expression(paste("Dam Volume per ", km^2))) + 
    theme(plot.title = element_text(hjust=0.5)) +
    coord_sf()

```


##Import data
###Dam volumes
```{r}
vol_by_grid <- readRDS("vol_by_grid.rds")
rvol <- rast(ncols = 174, nrows = 63, xmin = -120.985, xmax = -119.255, ymin = 37.325, ymax = 37.925)
rvol
values(rvol) <- vol_by_grid$total_vol_in_cell 
plot(rvol,1, colNA="grey70") #hmm upside down
#the raster is upside down so I flipped it. The final watershed should look like the one mapped below. 
rvol  <- flip(rvol, direction="vertical")
plot(rvol,1, colNA="grey70", main = "Dam Water Storage (m3)")
```

###Lis inputs 
```{r}
#lis inputs
inputs <- nc_open("lis_input_berkeley.nc")

```

###Pre-process LIS inputs so they make sense
```{r}
#Get latitude and longitude with the ncvar_get function and store each into their own object:
lat <- ncvar_get(inputs, "lat")
nlat <- dim(lat) #to check it matches the metadata: 63
lon <- ncvar_get(inputs, "lon")
nlon <- dim(lon) #to check, should be 174
# Check your lat lon dimensions match the information in the metadata we explored before:
print(c(nlon, nlat))

#Lat and Lon are both stored as two dimensional data arrays. Let's select just the unique lats and longs from those arrays: 
lat <- lat[1,]
lon <- lon[,1]
```

##Transpose inputs
```{r}
#The purpose of this code is to use the beaver dam volumes (the rvol raster from above) along with riverbed geometry (river height, width, length, floodplain height) for each grid cell to back-calculate the height a beaver dam would need to be to store the  appropriate volume of water. Then we gotta add that height to elevation to get total elevation of dam above sea level. 

#First, I pull arrays for individual variables from the netcdf file: 
HYMAP_grid_area <- ncvar_get(inputs, "HYMAP_grid_area")
HYMAP_floodplain_height <- ncvar_get(inputs, "HYMAP_floodplain_height")
HYMAP_river_width <- ncvar_get(inputs, "HYMAP_river_width")
HYMAP_river_length <- ncvar_get(inputs, "HYMAP_river_length")
HYMAP_river_height <- ncvar_get(inputs, "HYMAP_river_height")
elevation <- ncvar_get(inputs, "HYMAP_grid_elevation")

#Try transposing them:
HYMAP_grid_area_t <- t(HYMAP_grid_area)
HYMAP_floodplain_height_t <- aperm(HYMAP_floodplain_height, c(2,1,3)) #can't transpose this variable bc it's a 3d matrix
HYMAP_river_width_t <- t(HYMAP_river_width)
HYMAP_river_length_t <- t(HYMAP_river_length)
HYMAP_river_height_t <- t(HYMAP_river_height)
elevation_t <- t(elevation)

#Next, I rename my raster just for fun: 
damvol <- rvol
damvol.df <- as.matrix(t(raster(damvol)), xy=T) #Converted your raster to a matrix. Using the raster as-is is problematic as it uses the whole thing for every iteration of the loop instead of the specific cell you're interested in
damvol.dft <- t(damvol.df)
dim(damvol.dft)
str(damvol.dft)
```

```{r}
damheightsasl_array_t <- matrix(NA_real_, 63, 174) #Using a matrix seems easier than an array for some reason
floodheight_array_t <- matrix(NA_real_, 63, 174)
floodvol_array_t <- matrix(NA_real_, 63, 174)
rivervol_array_t <- matrix(NA_real_, 63, 174)
```


##For Loop
```{r}
for (i in 1:length(lat)){ #iterate over every longitude
  for (j in 1:length(lon)){ #iterate over every latitude
    damvol_val = damvol.dft[i,j] #Using the matrix instead of the raster in your original code
    rivervol = HYMAP_river_width_t[i,j] * HYMAP_river_length_t[i,j] * HYMAP_river_height_t[i,j] #calculate the volume of the river in that particular cell (width*length*height)
    floodvol = damvol_val + rivervol #calculate the volume of water left in the floodplain                             
    area = HYMAP_grid_area_t[i,j] #pull in area
    height = HYMAP_floodplain_height_t[i,j,] #pull in floodplain height, which is a 3 dim variable 
    volume = area * height #floodplane volume 
    model = lm(volume ~ height) #make a lm
    m = coef(model)[2]
    b = coef(model)[1]
    floodheight = (floodvol-b)/m #back calculate floodplain height
    floodheight[floodheight<0] = 0
    damheightsasl_array_t[i,j] = floodheight + elevation_t[i,j] #add to the elevation variable 
    floodheight_array_t[i,j] = floodheight
    floodvol_array_t[i,j] = floodvol
    rivervol_array_t[i,j] = rivervol
    
 }} 
```

###Viewing all the outputs
```{r}
e <- extent(c(-120.985, -119.255, 37.325, 37.925))

damheight_t <- rast(damheightsasl_array_t)
terra::set.ext(x = damheight_t, value = e)
terra::set.crs(damheight_t,wkt)
mapview(damheight_t)

floodheight_r <- rast(floodheight_array_t)
terra::set.ext(x = floodheight_r, value = e)
terra::set.crs(floodheight_r,wkt)
mapview(floodheight_r)

elevation_r <- rast(elevation_t)
terra::set.ext(x = elevation_r, value = e)
terra::set.crs(elevation_r,wkt)
mapview(elevation_r)
```

####Map
```{r}
ggplot() +
    geom_spatraster(data = floodheight_r, aes(fill = lyr.1)) +
    scale_fill_viridis(discrete = FALSE, option = "plasma", na.value = "grey70")+
    xlab(NULL)+
    ylab(NULL)+
       theme_minimal()+
    ggtitle(expression(paste("Dam heights (m)"))) + 
    theme(plot.title = element_text(hjust=0.5))

ggplot() +
    geom_spatraster(data = floodheight_r, aes(fill = lyr.1)) +
    scale_fill_gradient(low='#26A63A', high = '#FFD5E1')+
    xlab(NULL)+
    ylab(NULL)+
       theme_minimal()+
    ggtitle(expression(paste("Dam heights (m)"))) + 
    theme(plot.title = element_text(hjust=0.5)) 

```

```{r}
elevation_t[2,3] #this is 20m asl
floodheight_array_t[2,3] #this should NOT be NA
damheightsasl_array_t[1,2] #this should also NOT be NA 
#The mapviewer shows Floodheight_array to be 0.005 at [1,2]
#The mapviewer shows damheightsasl_array to be 30 at [1,2]

#Both sitautions are wrong. There should NOT be NAs at [1,2]. Also, there total damheights should be 20.005, not 30. What the fuck? 

floodheight_array_t[60,6]
elevation_t[60,6]
damheightsasl_array_t[60,6]
#That's correct now

#Okay now we have flipped things

elevation_t[2,3] #this is 16 asl
floodheight_array_t[2,3] #this should NOT be NA
damheightsasl_array_t[1,2] #this should also NOT be NA 


floodheight_array_t[60,6]
elevation_t[60,6]
damheightsasl_array_t[60,6]
#That's correct now
#But the issue is that it's indexed upside down for Python

```

###make NetCDF file
```{r}
damheight_spatrast <- rast(damheight_t)
writeCDF(damheight_spatrast,"damheight_elevation_6.nc",overwrite=TRUE)
```

###floodheight for Augusto 
```{r}
floodheight_spatrast <- rast(floodheight_array_t)
writeCDF(floodheight_spatrast,"floodheight_elevation_6.nc",overwrite=TRUE)
```