---
title: "AllSierras_LandUse"
output: html_document
---

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(dataRetrieval)
library(sf)
library(maps)
library(sp)
library(ncdf4)
library(sf)
library(sp)
library(lwgeom)
library(terra)
library(raster)
library(mapview)
library(tidyterra)
library(viridis)
library(MetBrewer)
library(stars)
library(data.table)
library(tigris)
library(patchwork)
library(car)
library(multcomp)
```

#Prep Work
###NASA projection
```{r}
wkt <- 'GEOGCS["WGS 84",DATUM["WGS_1984",SPHEROID["WGS 84",6378137,298.257223563,AUTHORITY["EPSG","7030"]],AUTHORITY["EPSG","6326"]],PRIMEM["Greenwich",0,AUTHORITY["EPSG","8901"]],UNIT["degree",0.0174532925199433,AUTHORITY["EPSG","9122"]],AXIS["Latitude",NORTH],AXIS["Longitude",EAST],AUTHORITY["EPSG","4326"]]'

#California Teal Albers EPSG = 3310
```

###Input Files (all together)
```{r}
AllSierras_DamCapacity <- readRDS("AllSierras_DamCapacity_volume.rds")
PADUS <- readRDS("PADUS.rds")
WatershedBoundaries <- readRDS("WatershedBoundaries.rds")
PADUS_clip = st_intersection(PADUS, WatershedBoundaries)
PADUS_Gap12 <- PADUS_clip[PADUS_clip$GAP_Sts==c(1,2),]
DamCap_GAP12 = st_intersection(AllSierras_DamCapacity, PADUS_Gap12)

```

###Dam Capacity
```{r}
# Read in the shapefile with st_read()
#AllSierras_DamCapacity <- st_read("C:/Users/jessie_moravek/Desktop/SierraNevadaBeavers/AllSierras_DamCapacity.shp")

#remove unneeded shapefiles
#file.remove(list.files(pattern = "AllSierras_DamCapacity*",recursive = F))

# Read in shapefile from previously imported shapefile
AllSierras_DamCapacity <- readRDS("AllSierras_DamCapacity_volume.rds")

#Project to Teal Albers
AllSierras_DamCapacity <- st_transform(AllSierras_DamCapacity, 3310)
saveRDS(AllSierras_DamCapacity,"AllSierras_DamCapacity.rds")
```

####Map
```{r}
palette_new <- colorRampPalette(colors = c("#ff0000", "#ffa500", "#ffff00", "#00FF50", "#4900FF"))(5)
#scales::show_col(palette_new)

ggplot(transform(AllSierras_DamCapacity, DamCapacity=cut(oCC_EX, breaks = c(0,.1,2,5,15,40), include.lowest=T))) +
    geom_sf(aes(color = DamCapacity)) +
    scale_color_manual(values=palette_new)+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Dam Capacity (Dams/km)"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) 

```

###PADUS
```{r}
# Read in the shapefile with st_read()
#PADUS <- st_read("C:/Users/jessie_moravek/Desktop/SierraNevadaBeavers/PADUS.shp")

#remove unneeded shapefiles
#file.remove(list.files(pattern = "PADUS*",recursive = F))

# Read in shapefile from previously imported shapefile
PADUS <- readRDS("PADUS.rds")

#Project to Teal Albers
PADUS <- st_transform(PADUS, 3310)
saveRDS(PADUS,"PADUS.rds")
```

###WatershedBoundaries
```{r}
# Read in the shapefile with st_read()
#WatershedBoundaries <- st_read("C:/Users/jessie_moravek/Desktop/SierraNevadaBeavers/WatershedBoundaries2.shp")

#remove unneeded shapefiles
#file.remove(list.files(pattern = "WatershedBoundaries*",recursive = F))

# Read in shapefile from previously imported shapefile
WatershedBoundaries <- readRDS("WatershedBoundaries.rds")

#Project to Teal Albers
WatershedBoundaries <- st_transform(WatershedBoundaries, 3310)
saveRDS(WatershedBoundaries,"WatershedBoundaries.rds")
```


###Clip PADUS to watershed boundary area
```{r}
#sf_use_s2(FALSE) #only needed if using wkt
PADUS_clip = st_intersection(PADUS, WatershedBoundaries)
```

####Map
```{r}
palette_discrete <- colorRampPalette(colors = c("#E41A1C", "#CCCCCC", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#0070FF"))(7)
#scales::show_col(palette_new)

ggplot() +
    geom_sf(data = PADUS_clip,aes(color = d_GAP_Sts)) +
    scale_fill_manual(values=palette_discrete)+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Restoration Limitations"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) 
```
###Select just GAP 1 and 2

```{r}
PADUS_Gap12 <- PADUS_clip[PADUS_clip$GAP_Sts==c(1,2),]
unique(PADUS_Gap12$FeatClass)
```

####Map
```{r}
palette_discrete <- colorRampPalette(colors = c("#E41A1C", "#CCCCCC", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#0070FF"))(7)
#scales::show_col(palette_new)

ggplot() +
    geom_sf(data = PADUS_Gap12,aes(color = d_GAP_Sts)) +
    scale_fill_manual(values=palette_discrete, aesthetics = "fill")+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Protected Areas (GAP 1 & 2)"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) 
```
###Clip BRAT to GAP 1 and 2 area
```{r}
DamCap_GAP12 = st_intersection(AllSierras_DamCapacity, PADUS_Gap12)

```

####Map
```{r}
palette_new <- colorRampPalette(colors = c("#ff0000", "#ffa500", "#ffff00", "#00FF50", "#4900FF"))(5)
#scales::show_col(palette_new)

ggplot(transform(DamCap_GAP12, DamCapacity=cut(oCC_EX, breaks = c(0,.1,2,5,15,40), include.lowest=T))) +
    geom_sf(aes(color = DamCapacity)) +
    scale_color_manual(values=palette_new)+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Dam Capacity (Dams/km)"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) 

```

###Percent Dam Capacity in Gap 1 and 2 PAs
```{r}
a <- sum(DamCap_GAP12$oCC_EX)
b <- sum(AllSierras_DamCapacity$oCC_EX)

percentDamCapacity_InGap <- a/b

#GAP12 areas hold ~25% of potential beaver dam building capacity
```
#All Dam Building Capacity
##Fee PAs
###Select just fee areas in gap 1 and 2

```{r}
PADUS_Fee <- PADUS_Gap12[PADUS_Gap12$FeatClass=="Fee",]

```
###Clip BRAT to fee
```{r}
DamCap_Fee = st_intersection(AllSierras_DamCapacity, PADUS_Fee)

```
###Percent Dam Capacity in fee
```{r}
c <- sum(DamCap_Fee$oCC_EX)
b <- sum(AllSierras_DamCapacity$oCC_EX)

percentDamCapacity_InFee <- c/b

#Fee areas hold ~25% of potential beaver dam building capacity
```

##Easements
###Select just Easements

```{r}
PADUS_Easements <- PADUS_Gap12[PADUS_Gap12$FeatClass=="Easement",]
```

###Clip BRAT to Easements
```{r}
DamCap_Easements = st_intersection(AllSierras_DamCapacity, PADUS_Easements)
```

###Percent Dam Capacity in Easements
```{r}
d <- sum(DamCap_Easements$oCC_EX)
e <- sum(AllSierras_DamCapacity$oCC_EX)

percentDamCapacity_InEasements <- d/e

#Easements areas hold ~25% of potential beaver dam building capacity
```

##Designations PAs
###Select just Designation areas in gap 1 and 2

```{r}
PADUS_Desig <- PADUS_Gap12[PADUS_Gap12$FeatClass=="Designation",]

```
###Clip BRAT to Designations
```{r}
DamCap_Desig = st_intersection(AllSierras_DamCapacity, PADUS_Desig)

```
###Percent Dam Capacity in Designations
```{r}
f <- sum(DamCap_Desig$oCC_EX)
g <- sum(AllSierras_DamCapacity$oCC_EX)

percentDamCapacity_InDesig <- f/g

#Designation areas hold ~25% of potential beaver dam building capacity
```


#Good Beaver Habitat
##Select just Good beaver habitat from BRAT
```{r}
#Just GAP 12 PADUS: PADUS_Gap12
DamCap_Blue <- AllSierras_DamCapacity[AllSierras_DamCapacity$oCC_EX > 15,]

```

##Clip BRAT to GAP 12
```{r}
DamCap_Blue_GAP12 <- st_intersection(DamCap_Blue, PADUS_Gap12)

```

###Percent good dam cap in GAP 12
```{r}
h <- sum(DamCap_Blue_GAP12$oCC_EX)
i <- sum(DamCap_Blue$oCC_EX)


percentDamCapacity_Blue_InGAP12 <-h/i

#Easements areas hold ~25% of potential beaver dam building capacity
```

##Fee PAs
###Select just fee areas in gap 1 and 2

```{r}
#PADUS_Fee <- PADUS_Gap12[PADUS_Gap12$FeatClass=="Fee",]

```
###Clip BRAT to fee
```{r}
DamCap_Blue_Fee = st_intersection(DamCap_Blue_GAP12, PADUS_Fee)

```
###Percent Dam Capacity in fee
```{r}
j <- sum(DamCap_Blue_Fee$oCC_EX)
k <- sum(DamCap_Blue_GAP12$oCC_EX)

percentDamCapacity_Blue_InFee <- j/k

#Fee areas hold ~25% of potential beaver dam building capacity
```

##Easements
###Select just Easements

```{r}
#PADUS_Easements <- PADUS_Gap12[PADUS_Gap12$FeatClass=="Easement",]
```

###Clip BRAT to Easements
```{r}
DamCap_Blue_Easements = st_intersection(DamCap_Blue_GAP12, PADUS_Easements)
```

###Percent Dam Capacity in Easements
```{r}
l <- sum(DamCap_Blue_Easements$oCC_EX)
m <- sum(DamCap_Blue_GAP12$oCC_EX)

percentDamCapacity_Blue_InEasements <- l/m

#Easements areas hold ~25% of potential beaver dam building capacity
```

##Designations PAs
###Select just Designation areas in gap 1 and 2

```{r}
#PADUS_Desig <- PADUS_Gap12[PADUS_Gap12$FeatClass=="Designation",]

```
###Clip BRAT to Designations
```{r}
DamCap_Blue_Desig = st_intersection(DamCap_Blue_GAP12, PADUS_Desig)

```
###Percent Dam Capacity in Designations
```{r}
n <- sum(DamCap_Blue_Desig$oCC_EX)
o <- sum(DamCap_Blue_GAP12$oCC_EX)

percentDamCapacity_Blue_InDesig <- n/o

#Designation areas hold ~25% of potential beaver dam building capacity
```
#Restoration Opportunities
##Select just highest rest opportunities from BRAT
```{r}
#Just GAP 12 PADUS: PADUS_Gap12
DamCap_RestOpp <- AllSierras_DamCapacity[AllSierras_DamCapacity$oPBRC_CR != "NA",]

```

##Clip BRAT to GAP 12
```{r}
DamCap_RestOpp_GAP12 <- st_intersection(DamCap_RestOpp, PADUS_Gap12)

```

##Percent Dam Capacity in Easements
```{r}
p <- sum(DamCap_RestOpp_GAP12$oCC_EX)
q <- sum(DamCap_RestOpp$oCC_EX)


percentDamCapacity_RestOpp_GAP12 <-p/q

#Easements areas hold ~25% of potential beaver dam building capacity
```
##Fee PAs
###Select just fee areas in gap 1 and 2

```{r}
#PADUS_Fee <- PADUS_Gap12[PADUS_Gap12$FeatClass=="Fee",]

```
###Clip BRAT to fee
```{r}
DamCap_RestOpp_Fee = st_intersection(DamCap_RestOpp_GAP12, PADUS_Fee)

```
###Percent Dam Capacity in fee
```{r}
r <- sum(DamCap_RestOpp_Fee$oCC_EX)
s <- sum(DamCap_RestOpp_GAP12$oCC_EX)

percentDamCapacity_RestOpp_InFee <- r/s

#Fee areas hold ~25% of potential beaver dam building capacity
```

##Easements
###Select just Easements

```{r}
#PADUS_Easements <- PADUS_Gap12[PADUS_Gap12$FeatClass=="Easement",]
```

###Clip BRAT to Easements
```{r}
DamCap_RestOpp_Easements = st_intersection(DamCap_RestOpp_GAP12, PADUS_Easements)
```

###Percent Dam Capacity in Easements
```{r}
t <- sum(DamCap_RestOpp_Easements$oCC_EX)
u <- sum(DamCap_RestOpp_GAP12$oCC_EX)

percentDamCapacity_RestOpp_InEasements <- t/u

#Easements areas hold ~25% of potential beaver dam building capacity
```

##Designations PAs
###Select just Designation areas in gap 1 and 2

```{r}
#PADUS_Desig <- PADUS_Gap12[PADUS_Gap12$FeatClass=="Designation",]

```
###Clip BRAT to Designations
```{r}
DamCap_RestOpp_Desig = st_intersection(DamCap_RestOpp_GAP12, PADUS_Desig)

```
###Percent Dam Capacity in Designations
```{r}
v <- sum(DamCap_RestOpp_Desig$oCC_EX)
w <- sum(DamCap_RestOpp_GAP12$oCC_EX)

percentDamCapacity_RestOpp_InDesig <- v/w

#Designation areas hold ~25% of potential beaver dam building capacity
```

#Figures
##inputs
```{r}
#PADUS_Gap12
#PADUS_Fee
#PADUS_Easements
#PADUS_Desig

background <- states(cb = FALSE)

background <- st_make_valid(background) %>%
  dplyr::mutate(valid =st_is_valid(background)) %>% 
  dplyr::filter(valid=="TRUE") 
#^Millie's validation code...use later

background <- background %>% filter(NAME == "California" | NAME == "Nevada")

background <- st_transform(background, 3310)

box <- st_bbox(AllSierras_DamCapacity)
background_crop <- st_crop(background, box)
```

##Palette
```{r}
#display_all(3, sequential=FALSE, colorblind_only=TRUE, direction=1, override.order=FALSE )
mycolors <-met.brewer(name ='Archambault',3, type="discrete")
mycolors[3]

mycolors <-met.brewer(name ='Archambault',4, type="discrete")
mycolors[3]

contcolors <- met.brewer(name = "Archambault", type = "continuous")
```
#Maps
##GAP 1&2
```{r}
padus <- ggplot() + 
  geom_sf(data = background, fill = "#dcdcdc", lwd = 0.1)+
  geom_sf(data = PADUS_Desig, fill = "#88a0dc", color = NA, aes(shape = "Designation")) +
  geom_sf(data = PADUS_Fee, fill = "#ed968c",color = NA, alpha = 0.8, aes(shape = "Fee")) +
  geom_sf(data = PADUS_Easements, fill = "#f9d14a", color = NA, aes(shape = "Easement")) +
  coord_sf(crs = st_crs(3310))+ 
  theme_minimal() +
  ggtitle(expression(paste("Protected Areas (GAP 1 & 2)"))) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  #coord_sf(xlim = c(-124.0, -116.8), ylim = c(34, 41)) +
  theme(panel.border = element_rect(colour = "black", fill=NA))

```

##Total Dam Capacity
```{r}
palette_new <- colorRampPalette(colors = c("#ff0000", "#ffa500", "#ffff00", "#00FF50", "#4900FF"))(5)
#scales::show_col(palette_new)

damcap <- ggplot() +
    geom_sf(data = background, fill = "#dcdcdc", lwd = 0.1)+
    geom_sf(data = transform(AllSierras_DamCapacity, DamCapacity=cut(oCC_EX, breaks = c(0,.1,2,5,15,40), include.lowest=T)),aes(color = DamCapacity)) +
    scale_color_manual(values=palette_new)+
    theme_minimal()+
    ggtitle(expression(paste("Maximum Dam Building Capacity"))) +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    theme(element_text(size=10),legend.position = c(.15,.2))+
    labs(color = "dams/km", size=10)+
    theme(panel.border = element_rect(colour = "black", fill=NA))

```
##Water Volume
```{r}
library(viridis)
vol <- ggplot() +
    geom_sf(data = background, fill = "#dcdcdc", lwd = 0.1)+
    geom_sf(data = AllSierras_DamCapacity, aes(color = volume)) +
    scale_color_viridis(discrete = FALSE, option = "plasma", na.value = "grey70")+
    theme_minimal()+
    ggtitle(expression(paste("Potential Dam Water Storage "))) +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    theme(element_text(size=10),legend.position ='right')+
    labs(color = "Water Storage (m3)", size=10)+
    theme(panel.border = element_rect(colour = "black", fill=NA))

```
####Map
```{r}
ggplot() +
    geom_sf(data = AllSierras_DamCapacity, aes(color = volume)) +
    scale_color_viridis(discrete = FALSE, option = "plasma", na.value = "grey70")+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Potential Water Storage (m3)"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) 

```


```{r}
(padus + damcap)
```


#Dam Capacity Boxplot
```{r}
test <- st_join(AllSierras_DamCapacity, PADUS_Gap12)
```
##Ttest GAP and other
```{r}
test$Protected <- with(test, ifelse(is.na(test$GAP_Sts), "Other", "GAP 1 & 2"))
test$Protected <- as.factor(test$Protected)
                                    
ggplot(data = test, aes(x=Protected, y = log(oCC_EX)))+
         geom_boxplot()+
         theme_minimal()
       
leveneTest(oCC_EX~Protected, data = test)#significant
ttest_1 <- t.test(oCC_EX ~ Protected, data = test)
summary(ttest_1)
p_value_1 <- ttest_1$p.value
t_value_1 <- ttest_1$statistic
estimate_gap <- ttest_1$estimate[1]
estimate_other <- ttest_1$estimate[2]
deg_freedom <- ttest_1$parameter
```

```{r}
gap <- subset(test, Protected == "GAP 1 & 2")
other <- subset(test, Protected == "Other")

p1 <- ggplot()+
  geom_boxplot(data = gap, aes(x=Protected, y = sqrt(oCC_EX)), fill = "#e78429")+
  geom_boxplot(data = other, aes(x=Protected, y = sqrt(oCC_EX)), fill= "grey60")+
  theme_minimal() +
  ggtitle(expression(paste("Dam Capacity"))) + 
  xlab("Protection Status")+
  ylab("Sqrt Dam Capacity (dams/km)")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))

```


##ANOVA within GAP 12
```{r}
DamCap_GAP12$FeatClass <- as.factor(DamCap_GAP12$FeatClass)
                                    
ggplot(data = DamCap_GAP12, aes(x=FeatClass, y = (oCC_EX)))+
         geom_boxplot()+
         theme_minimal()
       
leveneTest(oCC_EX~FeatClass, data = DamCap_GAP12)#significant
anova1 <- aov(oCC_EX ~ FeatClass, data = DamCap_GAP12)
anova1_residuals <- residuals(object = anova1)

ggplot(anova1, aes(x = .fitted, y = .resid))+
  geom_point() + theme_classic()
#shapiro.test(x = anova1_residuals)#too big

#try square root transformation
DamCap_GAP12 <- DamCap_GAP12 %>% mutate(sqrtvalue=sqrt(oCC_EX))
anova2 <- aov(sqrtvalue ~ FeatClass, data = DamCap_GAP12)

ggplot(data = DamCap_GAP12, aes(x=FeatClass, y = (sqrtvalue)))+
         geom_boxplot()+
         theme_minimal()

leveneTest(sqrtvalue~FeatClass, data = DamCap_GAP12)#significant

kruskal.test(sqrtvalue ~ FeatClass, data = DamCap_GAP12)

#PostHoc on the square root transformation
summary(anova2)
tukey <- TukeyHSD(anova2)
print(tukey)
```

```{r}
desig <- subset(DamCap_GAP12, FeatClass == "Designation")
fee <- subset(DamCap_GAP12, FeatClass == "Fee")
easement <- subset(DamCap_GAP12, FeatClass == "Easement")


p2 <- ggplot()+
  geom_boxplot(data = desig, aes(x=FeatClass, y = sqrt(oCC_EX)), fill = "#88a0dc") +
  geom_boxplot(data = fee, aes(x=FeatClass, y = sqrt(oCC_EX)), fill = "#ed968c") +
  geom_boxplot(data = easement, aes(x=FeatClass, y = sqrt(oCC_EX)), fill = "#f9d14a") +
  theme_minimal() +
  ggtitle(expression(paste("Dam Capacity"))) + 
  xlab("Land Ownership Type")+
  ylab("Sqrt Dam Capacity (dams/km)")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))

```

##Total dams GAP12 vs. other
```{r}
#Use mCC_EX_CT- maximum number of dams
gap <- subset(test, Protected == "GAP 1 & 2")
other <- subset(test, Protected == "Other")

p3 <- ggplot()+
  geom_col(data = gap, aes(x=Protected, y = mCC_EX_CT), fill = "#e78429")+
  geom_col(data = other, aes(x=Protected, y = mCC_EX_CT), fill= "grey60")+
  theme_minimal() +
  ggtitle(expression(paste("Number of Dams"))) + 
  xlab("Protection Status")+
  ylab("Number of Dams")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))
```
##Dams in GAP12 in Desig, Fee, and Easement
```{r}
desig <- subset(DamCap_GAP12, FeatClass == "Designation")
fee <- subset(DamCap_GAP12, FeatClass == "Fee")
easement <- subset(DamCap_GAP12, FeatClass == "Easement")


p4 <- ggplot()+
  geom_col(data = desig, aes(x=FeatClass, y = (mCC_EX_CT)), fill = "#88a0dc") +
  geom_col(data = fee, aes(x=FeatClass, y = (mCC_EX_CT)), fill = "#ed968c") +
  geom_col(data = easement, aes(x=FeatClass, y = (mCC_EX_CT)), fill = "#f9d14a") +
  theme_minimal() +
  ggtitle(expression(paste("Number of Dams"))) + 
  xlab("Land Ownership Type")+
  ylab("Number of Dams")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  #labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))
```

```{r}
(p3 + p1)/ (p4 + p2)
```




#High Quality Haibtat Boxplot
```{r}
#test <- st_join(AllSierras_DamCapacity, PADUS_Gap12)
highquality <- subset(test, oCC_EX >= 10) #pull out good habitat

```
##Ttest GAP and other
```{r}
highquality$Protected <- as.factor(highquality$Protected)
                                    
ggplot(data = highquality, aes(x=Protected, y = oCC_EX))+
         geom_boxplot()+
         theme_minimal()
       
leveneTest(oCC_EX~Protected, data = highquality)#significant
ttest_2 <- t.test(oCC_EX ~ Protected, data = highquality)
summary(ttest_2)
p_value_2 <- ttest_2$p.value
t_value_2 <- ttest_2$statistic
estimate_gap <- ttest_2$estimate[1]
estimate_other <- ttest_2$estimate[2]
deg_freedom <- ttest_2$parameter
```
####Map
```{r}
palette_new <- colorRampPalette(colors = c( "#00FF50", "#4900FF"))(2)
#scales::show_col(palette_new)

ggplot(transform(highquality, DamCapacity=cut(oCC_EX, breaks = c(0,.1,2,5,15,40), include.lowest=T))) +
    geom_sf(aes(color = DamCapacity)) +
    scale_color_manual(values=palette_new)+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Dam Capacity (Dams/km)"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) 

```

```{r}
gap2 <- subset(highquality, Protected == "GAP 1 & 2")
other2 <- subset(highquality, Protected == "Other")

h1 <- ggplot()+
  geom_boxplot(data = gap2, aes(x=Protected, y = sqrt(oCC_EX)), fill = "#e78429")+
  geom_boxplot(data = other2, aes(x=Protected, y = sqrt(oCC_EX)), fill= "grey60")+
  theme_minimal() +
  ggtitle(expression(paste("High Quality Dam Capacity in Protected and Non-Protected Areas"))) + 
  xlab("Protection Status")+
  ylab("Square Root of Dam Capacity (dams/km)")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))

```


##Ttest GAP and other
```{r}
highquality_gap12 <- subset(DamCap_GAP12, oCC_EX >= 10) #pull out good habitat
highquality_gap12$FeatClass <- as.factor(highquality_gap12$FeatClass)

                                    
ggplot(data = highquality_gap12, aes(x=FeatClass, y = (oCC_EX)))+
         geom_boxplot()+
         theme_minimal()
       
leveneTest(oCC_EX~FeatClass, data = highquality_gap12)#significant
anova1 <- aov(oCC_EX ~ FeatClass, data = highquality_gap12)
anova1_residuals <- residuals(object = anova1)

ggplot(anova1, aes(x = .fitted, y = .resid))+
  geom_point() + theme_classic()
#shapiro.test(x = anova1_residuals)#too big

#try square root transformation
highquality_gap12 <- highquality_gap12 %>% mutate(sqrtvalue=sqrt(oCC_EX))
anova2 <- aov(sqrtvalue ~ FeatClass, data = highquality_gap12)

ggplot(data = highquality_gap12, aes(x=FeatClass, y = (sqrtvalue)))+
         geom_boxplot()+
         theme_minimal()

leveneTest(sqrtvalue~FeatClass, data = highquality_gap12)#significant

kruskal.test(oCC_EX ~ FeatClass, data = highquality_gap12)#Hmm...not significant if nonparametric

#PostHoc on the square root transformation
summary(anova2)
tukey <- TukeyHSD(anova2)
print(tukey)
```

```{r}
desig <- subset(highquality_gap12, FeatClass == "Designation")
fee <- subset(highquality_gap12, FeatClass == "Fee")
easement <- subset(highquality_gap12, FeatClass == "Easement")


h2 <- ggplot()+
  geom_boxplot(data = desig, aes(x=FeatClass, y = sqrt(oCC_EX)), fill = "#88a0dc") +
  geom_boxplot(data = fee, aes(x=FeatClass, y = sqrt(oCC_EX)), fill = "#ed968c") +
  geom_boxplot(data = easement, aes(x=FeatClass, y = sqrt(oCC_EX)), fill = "#f9d14a") +
  theme_minimal() +
  ggtitle(expression(paste("High Quality Dam Capacity in GAP 1&2 Areas"))) + 
  xlab("Land Ownership Type")+
  ylab("Square Root of Dam Capacity (dams/km)")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))

```
##Total dams GAP12 vs. other
```{r}
#Use mCC_EX_CT- maximum number of dams
gap <- subset(highquality, Protected == "GAP 1 & 2")
other <- subset(highquality, Protected == "Other")

h3 <- ggplot()+
  geom_col(data = gap, aes(x=Protected, y = mCC_EX_CT), fill = "#e78429")+
  geom_col(data = other, aes(x=Protected, y = mCC_EX_CT), fill= "grey60")+
  theme_minimal() +
  ggtitle(expression(paste("Number of Dams"))) + 
  xlab("Protection Status")+
  ylab("Number of Dams")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))
```
##Dams in GAP12 in Desig, Fee, and Easement
```{r}
desig <- subset(highquality_gap12, FeatClass == "Designation")
fee <- subset(highquality_gap12, FeatClass == "Fee")
easement <- subset(highquality_gap12, FeatClass == "Easement")


h4 <- ggplot()+
  geom_col(data = desig, aes(x=FeatClass, y = (mCC_EX_CT)), fill = "#88a0dc") +
  geom_col(data = fee, aes(x=FeatClass, y = (mCC_EX_CT)), fill = "#ed968c") +
  geom_col(data = easement, aes(x=FeatClass, y = (mCC_EX_CT)), fill = "#f9d14a") +
  theme_minimal() +
  ggtitle(expression(paste("Number of Dams"))) + 
  xlab("Land Ownership Type")+
  ylab("Number of Dams")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  #labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))
```

```{r}
(h3 + h1) / (h4 + h2)
```


#Volume Boxplot
```{r}
test <- st_join(AllSierras_DamCapacity, PADUS_Gap12)
```
##Ttest GAP and other
```{r}
test$Protected <- with(test, ifelse(is.na(test$GAP_Sts), "Other", "GAP 1 & 2"))
test$Protected <- as.factor(test$Protected)
                                    
ggplot(data = test, aes(x=Protected, y = log(volume)))+
         geom_boxplot()+
         theme_minimal()
       
leveneTest(volume~Protected, data = test)#significant
ttest3 <- t.test(volume ~ Protected, data = test)
summary(ttest3)
p_value3 <- ttest3$p.value
t_value3 <- ttest3$statistic
estimate_gap3 <- ttest3$estimate[1]
estimate_other3 <- ttest3$estimate[2]
deg_freedom3 <- ttest3$parameter
```
```{r}
gap <- subset(test, Protected == "GAP 1 & 2")
other <- subset(test, Protected == "Other")

v1 <- ggplot()+
  geom_boxplot(data = gap, aes(x=Protected, y = sqrt(volume)), fill = "#e78429")+
  geom_boxplot(data = other, aes(x=Protected, y = sqrt(volume)), fill= "grey60")+
  theme_minimal() +
  ggtitle(expression(paste("Volume by Reach"))) + 
  xlab("Protection Status")+
  ylab("Square Root of Volume (m3)")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))

```

##ANOVA within GAP 12
```{r}
DamCap_GAP12$FeatClass <- as.factor(DamCap_GAP12$FeatClass)
                                    
ggplot(data = DamCap_GAP12, aes(x=FeatClass, y = (oCC_EX)))+
         geom_boxplot()+
         theme_minimal()
       
leveneTest(volume~FeatClass, data = DamCap_GAP12)#significant
anova1 <- aov(volume ~ FeatClass, data = DamCap_GAP12)
anova1_residuals <- residuals(object = anova1)

ggplot(anova1, aes(x = .fitted, y = .resid))+
  geom_point() + theme_classic()
#shapiro.test(x = anova1_residuals)#too big

#try square root transformation
DamCap_GAP12 <- DamCap_GAP12 %>% mutate(sqrtvalue=sqrt(volume))
anova2 <- aov(sqrtvalue ~ FeatClass, data = DamCap_GAP12)

ggplot(data = DamCap_GAP12, aes(x=FeatClass, y = (sqrtvalue)))+
         geom_boxplot()+
         theme_minimal()

leveneTest(sqrtvalue~FeatClass, data = DamCap_GAP12)#significant

kruskal.test(sqrtvalue ~ FeatClass, data = DamCap_GAP12)

#PostHoc on the square root transformation
summary(anova2)
tukey <- TukeyHSD(anova2)
print(tukey)
```

```{r}
desig <- subset(DamCap_GAP12, FeatClass == "Designation")
fee <- subset(DamCap_GAP12, FeatClass == "Fee")
easement <- subset(DamCap_GAP12, FeatClass == "Easement")


v2 <- ggplot()+
  geom_boxplot(data = desig, aes(x=FeatClass, y = sqrt(volume)), fill = "#88a0dc") +
  geom_boxplot(data = fee, aes(x=FeatClass, y = sqrt(volume)), fill = "#ed968c") +
  geom_boxplot(data = easement, aes(x=FeatClass, y = sqrt(volume)), fill = "#f9d14a") +
  theme_minimal() +
  ggtitle(expression(paste("Volume by Reach"))) + 
  xlab("Land Ownership Type")+
  ylab("Square Root of Volume (m3)")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))

```
##Total dams GAP12 vs. other
```{r}
#Use mCC_EX_CT- maximum number of dams
gap <- subset(test, Protected == "GAP 1 & 2")
other <- subset(test, Protected == "Other")

v3 <- ggplot()+
  geom_col(data = gap, aes(x=Protected, y = volume), fill = "#e78429")+
  geom_col(data = other, aes(x=Protected, y = volume), fill= "grey60")+
  theme_minimal() +
  ggtitle(expression(paste("Total Volume"))) + 
  xlab("Protection Status")+
  ylab("Total Volume (m3)")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))
```
##Dams in GAP12 in Desig, Fee, and Easement
```{r}
desig <- subset(DamCap_GAP12, FeatClass == "Designation")
fee <- subset(DamCap_GAP12, FeatClass == "Fee")
easement <- subset(DamCap_GAP12, FeatClass == "Easement")


v4 <- ggplot()+
  geom_col(data = desig, aes(x=FeatClass, y = (volume)), fill = "#88a0dc") +
  geom_col(data = fee, aes(x=FeatClass, y = (volume)), fill = "#ed968c") +
  geom_col(data = easement, aes(x=FeatClass, y = (volume)), fill = "#f9d14a") +
  theme_minimal() +
  ggtitle(expression(paste("Total Volume"))) + 
  xlab("Land Ownership Type")+
  ylab("Total Volume (m3)")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  #labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))
```

```{r}
(v3 + v1) / (v4 + v2)
```



#High Quality Volume Boxplot
```{r}
#test <- st_join(AllSierras_DamCapacity, PADUS_Gap12)
highquality <- subset(test, oCC_EX >= 10) #pull out good habitat

```

####Map
```{r}
ggplot() +
    geom_sf(data = highquality, aes(color = volume)) +
    scale_color_viridis(discrete = FALSE, option = "plasma", na.value = "grey70")+
    xlab(NULL)+
    ylab(NULL)+
    ggtitle(expression(paste("Potential Water Storage (m3)"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) 

```
##Ttest GAP and other
```{r}
highquality$Protected <- with(highquality, ifelse(is.na(highquality$GAP_Sts), "Other", "GAP 1 & 2"))
highquality$Protected <- as.factor(highquality$Protected)
                                    
ggplot(data = highquality, aes(x=Protected, y = log(volume)))+
         geom_boxplot()+
         theme_minimal()
       
leveneTest(volume~Protected, data = highquality)#significant
ttest3 <- t.test(volume ~ Protected, data = highquality)
summary(ttest3)
p_value3 <- ttest3$p.value
t_value3 <- ttest3$statistic
estimate_gap3 <- ttest3$estimate[1]
estimate_other3 <- ttest3$estimate[2]
deg_freedom3 <- ttest3$parameter
```


```{r}
gap <- subset(highquality, Protected == "GAP 1 & 2")
other <- subset(highquality, Protected == "Other")

hv1 <- ggplot()+
  geom_boxplot(data = gap, aes(x=Protected, y = sqrt(volume)), fill = "#e78429")+
  geom_boxplot(data = other, aes(x=Protected, y = sqrt(volume)), fill= "grey60")+
  theme_minimal() +
  ggtitle(expression(paste("High Quality Volume"))) + 
  xlab("Protection Status")+
  ylab("Square Root of Volume (m3)")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))

```
##ANOVA within GAP 12
```{r}
highquality_gap12 <- subset(DamCap_GAP12, oCC_EX >= 10) #pull out good habitat
highquality_gap12$FeatClass <- as.factor(highquality_gap12$FeatClass)

                                                 
ggplot(data = highquality_gap12, aes(x=FeatClass, y = (oCC_EX)))+
         geom_boxplot()+
         theme_minimal()
       
leveneTest(volume~FeatClass, data = highquality_gap12)#significant
anova1 <- aov(volume ~ FeatClass, data = highquality_gap12)
anova1_residuals <- residuals(object = anova1)

ggplot(anova1, aes(x = .fitted, y = .resid))+
  geom_point() + theme_classic()
#shapiro.test(x = anova1_residuals)#too big

#try square root transformation
highquality_gap12 <- highquality_gap12 %>% mutate(sqrtvalue=sqrt(volume))
anova2 <- aov(sqrtvalue ~ FeatClass, data = highquality_gap12)

ggplot(data = highquality_gap12, aes(x=FeatClass, y = (sqrtvalue)))+
         geom_boxplot()+
         theme_minimal()

leveneTest(sqrtvalue~FeatClass, data = highquality_gap12)#significant

kruskal.test(sqrtvalue ~ FeatClass, data = highquality_gap12)

#PostHoc on the square root transformation
summary(anova2)
tukey <- TukeyHSD(anova2)
print(tukey)
```


```{r}
desig <- subset(highquality_gap12, FeatClass == "Designation")
fee <- subset(highquality_gap12, FeatClass == "Fee")
easement <- subset(highquality_gap12, FeatClass == "Easement")


hv2 <- ggplot()+
  geom_boxplot(data = desig, aes(x=FeatClass, y = sqrt(volume)), fill = "#88a0dc") +
  geom_boxplot(data = fee, aes(x=FeatClass, y = sqrt(volume)), fill = "#ed968c") +
  geom_boxplot(data = easement, aes(x=FeatClass, y = sqrt(volume)), fill = "#f9d14a") +
  theme_minimal() +
  ggtitle(expression(paste("High Quality Volume in GAP 1&2 Areas"))) + 
  xlab("Land Ownership Type")+
  ylab("Square Root of Volume (m3)")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))

```
##Total dams GAP12 vs. other
```{r}
#Use mCC_EX_CT- maximum number of dams
gap <- subset(highquality, Protected == "GAP 1 & 2")
other <- subset(highquality, Protected == "Other")

hv3 <- ggplot()+
  geom_col(data = gap, aes(x=Protected, y = volume), fill = "#e78429")+
  geom_col(data = other, aes(x=Protected, y = volume), fill= "grey60")+
  theme_minimal() +
  ggtitle(expression(paste("High Quality Volume"))) + 
  xlab("Protection Status")+
  ylab("Volume (m3)")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))
```
##Dams in GAP12 in Desig, Fee, and Easement
```{r}
desig <- subset(highquality_gap12, FeatClass == "Designation")
fee <- subset(highquality_gap12, FeatClass == "Fee")
easement <- subset(highquality_gap12, FeatClass == "Easement")


hv4 <- ggplot()+
  geom_col(data = desig, aes(x=FeatClass, y = (volume)), fill = "#88a0dc") +
  geom_col(data = fee, aes(x=FeatClass, y = (volume)), fill = "#ed968c") +
  geom_col(data = easement, aes(x=FeatClass, y = (volume)), fill = "#f9d14a") +
  theme_minimal() +
  ggtitle(expression(paste("High Quality Volume in GAP 1&2 Areas"))) + 
  xlab("Land Ownership Type")+
  ylab("Volume (m3)")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  #labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))
```

```{r}
(hv3 + hv1) / (hv4 + hv2)
```


#Bar Plots
##Total dams GAP12 vs. other
```{r}
#Use mCC_EX_CT- maximum number of dams
gap <- subset(test, Protected == "GAP 1 & 2")
other <- subset(test, Protected == "Other")

ggplot()+
  geom_col(data = gap, aes(x=Protected, y = mCC_EX_CT), fill = "#e78429")+
  geom_col(data = other, aes(x=Protected, y = mCC_EX_CT), fill= "grey60")+
  theme_minimal() +
  ggtitle(expression(paste("Number of Dams in Protected and Non-Protected Areas"))) + 
  xlab("Protection Status")+
  ylab("Number of Dams")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))
```
##Dams in GAP12 in Desig, Fee, and Easement
```{r}
desig <- subset(DamCap_GAP12, FeatClass == "Designation")
fee <- subset(DamCap_GAP12, FeatClass == "Fee")
easement <- subset(DamCap_GAP12, FeatClass == "Easement")


ggplot()+
  geom_col(data = desig, aes(x=FeatClass, y = (mCC_EX_CT)), fill = "#88a0dc") +
  geom_col(data = fee, aes(x=FeatClass, y = (mCC_EX_CT)), fill = "#ed968c") +
  geom_col(data = easement, aes(x=FeatClass, y = (mCC_EX_CT)), fill = "#f9d14a") +
  theme_minimal() +
  ggtitle(expression(paste("Number of Dams in GAP 1&2 Areas"))) + 
  xlab("Land Ownership Type")+
  ylab("Number of Dams")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  #theme(legend.position = c(.15,.15))+
  #labs(shape = "")+
  theme(panel.border = element_rect(colour = "black", fill=NA))
```
##



