---
title: "UpperMerced_DamHeights_2"
output: html_document
date: "2024-01-11"
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(dataRetrieval)
library(ggsn)
library(sf)
library(maps)
library(sp)
library(ncdf4)
library(sf)
library(sp)
library(lwgeom)
library(terra)
library(raster)
library(plot.matrix)

```


##NASA projection
```{r}
wkt <- 'GEOGCS["WGS 84",DATUM["WGS_1984",SPHEROID["WGS 84",6378137,298.257223563,AUTHORITY["EPSG","7030"]],AUTHORITY["EPSG","6326"]],PRIMEM["Greenwich",0,AUTHORITY["EPSG","8901"]],UNIT["degree",0.0174532925199433,AUTHORITY["EPSG","9122"]],AXIS["Latitude",NORTH],AXIS["Longitude",EAST],AUTHORITY["EPSG","4326"]]'
```


##Dam Capacity
```{r}
# Read in the shapefile with st_read()
#damCapacity2 <- st_read("C:/Users/jessie_moravek/Desktop/SierraNevadaBeavers/DamCapacity2.shp")

#remove unneeded shapefiles
#file.remove(list.files(pattern = "damCapacity2*",recursive = F))

# Read in shapefile from previously imported shapefile
damCapacity2 <- readRDS("damCapacity2")

#Project to wkt (NASA projection)
damCapacity2 <- st_transform(damCapacity2, wkt)
saveRDS(damCapacity2,"damCapacity2")
```

##Make a grid
```{r}
#Make a bounding box extent
xmin <- -120.985
xmax <- -119.255
ymin <- 37.325
ymax <- 37.925
extent <- extent(xmin, xmax, ymin, ymax)

#Make a bounding box in the correct projection
study_area <- st_bbox(extent, crs = st_crs(wkt))

#Is wkt in metric or lat/long coordinate system? st_is_longlat gives TRUE, so it is NOT in metric.
st_is_longlat(study_area)

#Make a grid of the correct size
study_area = st_make_grid(study_area, n = c(174,63))

#Convert to an sf dataframe for future manipulation
study_area = st_as_sf(study_area)

#Add an ID column
study_area$id = 1:nrow(study_area)
```

##Volume calculation
```{r}
damCapacity2_v <- damCapacity2
for (i in damCapacity2_v){ #for each stream segment in the BRAT model output shapefile
  damCt <- damCapacity2_v$mCC_EX_CT #number of beaver dams per segment (calculated based on dam density and segment length as the column mCC_EX_CT in the attribute table)
  slope <- damCapacity2_v$iGeo_Slope #meters
  da <- damCapacity2_v$iGeo_DA #m2
  width <- ifelse(da < 4.95,2.18*(da)^0.191,
                  ifelse(da >= 4.95 & da < 377,1.41*(da)^0.462,7.18*(da)^0.183)) #based on Scamardo et al. 2022
  height <- 1 #assume all beaver dams are 1m tall
  length <- height/slope #trigonometry: slope = rise/run, so run = rise/slope
  maxLength<- (damCapacity2_v$iGeo_Len/damCt) #set a maximum length so that we can change volume calculations if there are so many dams that they all run together
  volume <- ifelse(length < maxLength, 0.5*(height)*(width)*(length)*(damCt),(height)*(width)*(maxLength)*(damCt))
  surfaceArea <- ifelse(length < maxLength, (width)*(length)*(damCt),(width)*(maxLength)*(damCt))
  damCapacity2_v$width <- width
  damCapacity2_v$volume <- volume #m3
  damCapacity2_v$surfaceArea <- surfaceArea #m2
} 


saveRDS(damCapacity2_v,"damCapacity2_v.rds")
```

##Volume by grid cell
```{r}
sf_use_s2(FALSE)
intersection <- st_intersection(study_area, damCapacity2_v) %>% 
  mutate(length = st_length(.)) %>% #creates a "length" column that is the length of that particular segment in the grid cell from the "id" column
  st_drop_geometry() # complicates things in joins later on

study_area <- study_area %>% 
  left_join(intersection, by = "id")

vol_by_grid <- study_area %>% 
  mutate(percent_len_in_cell = length/iGeo_Len) %>% 
  mutate(segment_vol_in_cell = percent_len_in_cell*volume) %>% 
  group_by(id) %>% summarize(total_vol_in_cell = sum(segment_vol_in_cell))

vol_by_grid$total_vol_in_cell <- as.numeric(vol_by_grid$total_vol_in_cell) #convert volume col to numeric

saveRDS(vol_by_grid,"vol_by_grid.rds")
vol_by_grid <- readRDS("vol_by_grid.rds")

```

###Map
```{r}

ggplot() +
    geom_sf(data = vol_by_grid, aes(fill = total_vol_in_cell), color = NA) +
    #geom_sf(data = damCapacity2, col = "blue") +
    scale_fill_gradient2(low = "gray70", high = "#078FF7") +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle( expression(paste("Volume per ", km^2))) + 
    theme_minimal() 
```
##Compare original volumes to new volumes

```{r}
summary(vol_by_grid$total_vol_in_cell)
sum(vol_by_grid$total_vol_in_cell,na.rm=T)
vol_by_grid %>% arrange(desc(total_vol_in_cell))

#Generally similar to original volumes in an old projection, if not slighlty higher
```

#Import data
##Dam volumes
```{r}
vol_by_grid <- readRDS("vol_by_grid.rds")

#volumes
rvol <- vol_by_grid

#make it a raster
rvol <- rast(ncols = 174, nrows = 63, xmin = -120.985, xmax = -119.255, ymin = 37.325, ymax = 37.925)
rvol
values(rvol) <- vol_by_grid$total_vol_in_cell 
plot(rvol,1, colNA="grey70") #hmm upside down
rvol  <- flip(rvol, direction="vertical")
plot(rvol,1, colNA="grey70")
```

##Lis inputs 
```{r}
#lis inputs
inputs <- nc_open("lis_input.nc")
print(inputs)

attributes(inputs$var)
attributes(inputs$dim)

```

##Organize Inputs data 
```{r}
# Get latitude and longitude with the ncvar_get function and store each into their own object:
lat <- ncvar_get(inputs, "lat")
nlat <- dim(lat) #to check it matches the metadata: 63
lon <- ncvar_get(inputs, "lon")
nlon <- dim(lon) #to check, should be 174
# Check your lat lon dimensions match the information in the metadata we explored before:
print(c(nlon, nlat))

#interesting they do not match. Lat and Lon are both stored as two dimensional data arrays. Let's select just the unique lats and longs from those arrays: 
lat <- lat[1,]
lon <- lon[,1]

#now pull an array, just to check
HYMAP_grid_area <- ncvar_get(inputs, "HYMAP_grid_area")
dim(HYMAP_grid_area) #to check; this should give you 174 and 63


elevation <- ncvar_get(inputs, "ELEVATION")
image(lon, lat, elevation)
summary(elevation)
```

#Dam volumes
```{r}
HYMAP_grid_area <- ncvar_get(inputs, "HYMAP_grid_area")
HYMAP_floodplain_height <- ncvar_get(inputs, "HYMAP_floodplain_height")
HYMAP_river_width <- ncvar_get(inputs, "HYMAP_river_width")
HYMAP_river_length <- ncvar_get(inputs, "HYMAP_river_length")
HYMAP_river_height <- ncvar_get(inputs, "HYMAP_river_height")
elevation <- ncvar_get(inputs, "ELEVATION")
damvol <- rvol

for (i in 1:length(lon)){
  for (j in 1:length(lat))
    damvol = damvol
    rivervol = HYMAP_river_width[i,j] * HYMAP_river_length[i,j] * HYMAP_river_height[i,j]
    floodvol = damvol - rivervol
    area = HYMAP_grid_area[i,j]
    height = HYMAP_floodplain_height[i,j,]
    volume = area * height
    model = lm(volume ~ height)
    m = coef(model)[2]
    b = coef(model)[1]
    floodheight = (floodvol-b)/m
    values(floodheight)[values(floodheight) < 0] = 0
    damheightsasl = floodheight + elevation[i,j]
    
    
    #totaldamheights = floodheight + HYMAP_river_height[i,j]
    #print(floodheight)
} 

plot(floodheight,1, colNA="grey70")
#plot(totaldamheights,1, colNA="grey70")
plot(damheightsasl,1, colNA="grey70")

#Attempting to add elevation highlights a weird problem-- it adds the SAME elevation to every cell. Why? When looking at the elevation nc slice, elevation differs by cell, as it should. But in this for loop, something goes wrong and elevation is the same in every cell. 
#Maybe this is going wrong because damvol is a raster, and all the other values are arrays (derived from slices of nc files?)
```

#Making rvol a .nc file
```{r}
library(RNetCDF)

writeCDF(rvol,"rvol.nc", overwrite=TRUE)

rvolnc <- nc_open("rvol.nc")
print(rvolnc)

# Get latitude and longitude with the ncvar_get function and store each into their own object:
  lat <- ncvar_get(rvolnc, "latitude")
  nlat <- dim(lat) #to check it matches the metadata: 63
  lon <- ncvar_get(rvolnc, "longitude")
  nlon <- dim(lon) #to check, should be 174
  # Check your lat lon dimensions match the information in the metadata we explored before:
  print(c(nlon, nlat))

attributes(rvolnc$var)
attributes(rvolnc$dim)

#now pull an array
total_vol_in_cell <- ncvar_get(rvolnc, "rvol")
dim(total_vol_in_cell) #to check; this should give you 174 and 63


```


```{r}
HYMAP_grid_area <- ncvar_get(inputs, "HYMAP_grid_area")
HYMAP_floodplain_height <- ncvar_get(inputs, "HYMAP_floodplain_height")
HYMAP_river_width <- ncvar_get(inputs, "HYMAP_river_width")
HYMAP_river_length <- ncvar_get(inputs, "HYMAP_river_length")
HYMAP_river_height <- ncvar_get(inputs, "HYMAP_river_height")
elevation <- ncvar_get(inputs, "ELEVATION")
damvol <- ncvar_get(rvolnc, "rvol")

damheightsasl_array <- array(dim = c(174,63))
for (i in 1:length(lon)){
  for (j in 1:length(lat)){
    damvol = damvol
    rivervol = HYMAP_river_width[i,j] * HYMAP_river_length[i,j] * HYMAP_river_height[i,j]
    floodvol = damvol - rivervol
    area = HYMAP_grid_area[i,j]
    height = HYMAP_floodplain_height[i,j,]
    volume = area * height
    model = lm(volume ~ height)
    m = coef(model)[2]
    b = coef(model)[1]
    floodheight = (floodvol-b)/m
    floodheight[floodheight < 0] = 0
    damheightsasl_array[i,j] = floodheight + elevation[i,j]
    #damheightsasl_array = damheightsasl[i,j]
 
    #totaldamheights = floodheight + HYMAP_river_height[i,j]
    #print(floodheight)
  }} 

damheightsasl_array
```

#Turn damheightsasl_array into a raster to make it plotable
```{r}
#make it a raster
damheightsasl_r <- rast(ncols = 174, nrows = 63, xmin = -120.985, xmax = -119.255, ymin = 37.325, ymax = 37.925)
damheightsasl_r

values(damheightsasl_r) <- t(damheightsasl_array)

plot(damheightsasl_r,1, colNA="grey70") #hmm upside down
damheightsasl_r  <- flip(damheightsasl_r, direction="vertical")
plot(damheightsasl_r,1, colNA="grey70")
```

#Turn elevation into a raster to make it plotable
```{r}
#make it a raster
elevation_r <- rast(ncols = 174, nrows = 63, xmin = -120.985, xmax = -119.255, ymin = 37.325, ymax = 37.925)
elevation_r

values(elevation_r) <- t(elevation)

plot(elevation_r,1, colNA="grey70") #hmm upside down
elevation_r  <- flip(elevation_r, direction="vertical")
plot(elevation_r,1, colNA="grey70")
```
#Huh...maybe damheights and elevation are the same? Need to double check
##Run for JUSt dam heights (sans elevation)

```{r}
HYMAP_grid_area <- ncvar_get(inputs, "HYMAP_grid_area")
HYMAP_floodplain_height <- ncvar_get(inputs, "HYMAP_floodplain_height")
HYMAP_river_width <- ncvar_get(inputs, "HYMAP_river_width")
HYMAP_river_length <- ncvar_get(inputs, "HYMAP_river_length")
HYMAP_river_height <- ncvar_get(inputs, "HYMAP_river_height")
elevation <- ncvar_get(inputs, "ELEVATION")
damvol <- ncvar_get(rvolnc, "rvol")

damheights_array <- array(dim = c(174,63))
for (i in 1:length(lon)){
  for (j in 1:length(lat)){
    damvol = damvol
    rivervol = HYMAP_river_width[i,j] * HYMAP_river_length[i,j] * HYMAP_river_height[i,j]
    floodvol = damvol - rivervol
    area = HYMAP_grid_area[i,j]
    height = HYMAP_floodplain_height[i,j,]
    volume = area * height
    model = lm(volume ~ height)
    m = coef(model)[2]
    b = coef(model)[1]
    floodheight = (floodvol-b)/m
    floodheight[floodheight < 0] = 0
    damheights_array[i,j] = floodheight[i,j] 
    #damheightsasl_array = damheightsasl[i,j]
 
    #totaldamheights = floodheight + HYMAP_river_height[i,j]
    #print(floodheight)
  }} 

damheights_array
max(damheights_array)
#Huh??? these are way too small. Does the dam heights variable change whether you do it by raster or by nc????
```

##Try with raster for the damvol variable
```{r}
HYMAP_grid_area <- ncvar_get(inputs, "HYMAP_grid_area")
HYMAP_floodplain_height <- ncvar_get(inputs, "HYMAP_floodplain_height")
HYMAP_river_width <- ncvar_get(inputs, "HYMAP_river_width")
HYMAP_river_length <- ncvar_get(inputs, "HYMAP_river_length")
HYMAP_river_height <- ncvar_get(inputs, "HYMAP_river_height")
elevation <- ncvar_get(inputs, "ELEVATION")
damvol <- rvol

damheights_array <- array(dim = c(174,63))
for (i in 1:length(lon)){
  for (j in 1:length(lat)){
    damvol = damvol
    rivervol = HYMAP_river_width[i,j] * HYMAP_river_length[i,j] * HYMAP_river_height[i,j]
    floodvol = damvol - rivervol
    area = HYMAP_grid_area[i,j]
    height = HYMAP_floodplain_height[i,j,]
    volume = area * height
    model = lm(volume ~ height)
    m = coef(model)[2]
    b = coef(model)[1]
    floodheight = (floodvol-b)/m
    floodheight[floodheight < 0] = 0
    damheights_array= floodheight 
    #damheightsasl_array = damheightsasl[i,j]
 
    #totaldamheights = floodheight + HYMAP_river_height[i,j]
    #print(floodheight)
  }} 

damheights_array
#okay that makes more sense...but why are the values different?
```
#F this. Would apply work better???

```{r}
sFunc <- function(x, y){
  if(is.na(x)){
    return(ds2[ds2[,"colA"] == y, "colB"])
  } else {
    return(x)
  }
}

mapply(sFunc, ds1$col1, ds1$col2)


mapply()
```

