---
title: "StreamGaugeQuery"
output: html_document
date: "2023-04-22"
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(dataRetrieval)
library(ggsn)
library(sf)
library(maps)

```
#USGS
#Search for relevant parameters
##NLDAS parameters
  
```{r}
#read in all the NWIS parameter codes 
pcode <- readNWISpCode("all")

#temperature codes
tempCodes <- pcode[grep("temperature",
                      pcode$parameter_nm,
                      ignore.case=TRUE),]

tempCodes
#Temperature, air, degrees Celsius = 00020

```
#Temperature
##Count UM gauges
```{r}
UM_gauges <- whatNWISsites(huc="18040008", 
                          parameterCd='00020')
nrow(UM_gauges)
#there are 72 possible gauges in the Upper Merced

names(UM_gauges)

UM_gauges
```

#Precipitation
##Count UM gauges
```{r}
#precipitation codes
precipCodes <- pcode[grep("precipitation",
                      pcode$parameter_nm,
                      ignore.case=TRUE),]

precipCodes
#Precipitation, mm = 99772
#Precipitation, cumulative, mm = 72194

precipCodesAll <- precipCodes[,"parameter_cd"]
```

```{r}
UM_gauges2 <- whatNWISsites(huc="18040008", 
                          parameterCd=precipCodesAll)

```
#Humidity
##Count UM Gauges
```{r}
humidCodes<- pcode[grep("humidity",
                      pcode$parameter_nm,
                      ignore.case=TRUE),]
humidCodes
#Relative humidity, % = 00052
```

```{r}
UM_gauges3 <- whatNWISsites(huc="18040008", 
                          parameterCd='00052')
#no humidity gauge

```

#Wind
##Count UM Gauges
```{r}
windCodes <- pcode[grep("wind",
                      pcode$parameter_nm,
                      ignore.case=TRUE),]

windCodes
#Wind speed (m/s) = 62625
```

```{r}
UM_gauges4 <- whatNWISsites(huc="18040008", 
                          parameterCd='00035')
#1 wind gauges
```

#Pressure
##Count UM Gauges
```{r}
pressureCodes <- pcode[grep("pressure",
                      pcode$parameter_nm,
                      ignore.case=TRUE),]

pressureCodes
#Barometric Pressure, millimeters of mercury = 00025
```

```{r}
UM_gauges5 <- whatNWISsites(huc="18040008", 
                          parameterCd='00025')
#84 surface pressure sites
```

#Radiation
##Count UM Gauges
```{r}
radCodes <- pcode[grep("radiation",
                      pcode$parameter_nm,
                      ignore.case=TRUE),]

radCodes
#Shortwave radiation, downward intensity, watts per square meter = 72186	
#Longwave radiation, downward intensity, watts per square meter	= 72175

radCodesAll <- radCodes[,"parameter_cd"]

```

```{r}
UM_gauges6 <- whatNWISsites(huc="18040008", 
                          parameterCd='72186')
#No SW raditation 

UM_gauges7 <- whatNWISsites(huc="18040008", 
                          parameterCd='72175')

#No LW radiation
```

#Evaporation
##Count UM Gauges
```{r}
evapCodes <- pcode[grep("evaporation",
                      pcode$parameter_nm,
                      ignore.case=TRUE),]

evapCodes
#Evaporation per recording interval, millimeters= 72200

evapCodesAll <- evapCodes[,"parameter_cd"]

```

```{r}
UM_gauges8 <- whatNWISsites(huc="18040008", 
                          parameterCd=evapCodesAll)
#No evap gauges
```

#Maps
##Get UM boundary shapefile
```{r}
getwd()

# Unzip the shapefile
#unzip("WBD_18_HU2_Shape.zip")

# Read in the shapefile with st_read()
shapefile <- st_read("C:/Users/jessi/Desktop/SierraNevadaBeavers/SierraNevadaBeavers/Shape/WBDHU8.shp")

# Select just the Upper Merced Watershed
UM_boundary <- st_write(shapefile[shapefile$huc8=="18040008",], "UM_boundary.shp", append = FALSE)

#remove unneeded shapefiles
file.remove(list.files(pattern = "WBD_18_HU2_Shape*",recursive = F))

#Project to California Teal Albers (NAD83) EPSG code 3310
UM_boundary <- st_transform(UM_boundary, 3310)
```
##UM Boundary Map
```{r}
ggplot() +
    geom_sf(data = UM_boundary) +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle("Upper Merced)") + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    north(sf_UM, symbol=10, location="bottomright") + 
    scalebar(UM_boundary,
             dist=10, dist_unit="km", st.size = 3,
             transform=FALSE)
```
##Transform gauge point data
```{r}
#Make the gauges point file into a sf object. Note that in the st_as_sf function, crs needs to the projection the file is ALREADY IN, not the projection you want. Specify the projection you want using st_transform
sf_UM <- st_as_sf(UM_gauges, 
                  coords = c("dec_long_va", "dec_lat_va"),
                  crs = 4269)

#Reproject gauge points
sf_UM <- st_transform(sf_UM, 3310)
```
##Transform gauge point data
```{r}
#Make the gauges point file into a sf object. Note that in the st_as_sf function, crs needs to the projection the file is ALREADY IN, not the projection you want. Specify the projection you want using st_transform
sf_UM4 <- st_as_sf(UM_gauges4, 
                  coords = c("dec_long_va", "dec_lat_va"),
                  crs = 4269)

#Reproject gauge points
sf_UM4 <- st_transform(sf_UM4, 3310)
```

##Transform gauge point data
```{r}
#Make the gauges point file into a sf object. Note that in the st_as_sf function, crs needs to the projection the file is ALREADY IN, not the projection you want. Specify the projection you want using st_transform
sf_UM5 <- st_as_sf(UM_gauges5, 
                  coords = c("dec_long_va", "dec_lat_va"),
                  crs = 4269)

#Reproject gauge points
sf_UM5 <- st_transform(sf_UM5, 3310)
```

##Map
```{r}
ggplot() +
    geom_sf(data = UM_boundary) +
    geom_sf(data = sf_UM5, color = "red", show.legend=T) + 
    geom_sf(data = sf_UM, color = "black", show.legend=T) +
    geom_sf(data = sf_UM4, color = "blue", show.legend=T) +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle("Barometric Pressure Gauge Locations (n = 84) \n Air Temperature Gauge Locations (n = 72) \n Wind Gauge Locations (n = 1)") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    north(sf_UM, symbol=10, location="bottomright") + 
    scalebar(UM_boundary,
             dist=10, dist_unit="km", st.size = 3,
             transform=FALSE)
```

#CDEC
##Import station data
```{r}
cdecStations <- read.csv("Station_Search_CDEC.csv")
```
##Transform location data
```{r}
#Make the gauges point file into a sf object. Note that in the st_as_sf function, crs needs to the projection the file is ALREADY IN, not the projection you want. Specify the projection you want using st_transform
cdecStations <- st_as_sf(cdecStations, 
                  coords = c("Longitude", "Latitude"),
                  crs = 4326)

#Reproject gauge points
cdecStations <- st_transform(cdecStations, 3310)
```
##Map
```{r}
ggplot() +
    geom_sf(data = UM_boundary) +
    geom_sf(data = cdecStations) + 
    xlab(NULL)+
    ylab(NULL)+
    ggtitle("CDEC stations") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    north(sf_UM, symbol=10, location="bottomright") + 
    scalebar(UM_boundary,
             dist=10, dist_unit="km", st.size = 3,
             transform=FALSE)
```
##Clip to UM boundary
```{r}
UM_cdec <- st_intersection(UM_boundary,cdecStations)

ggplot() +
    geom_sf(data = UM_boundary) +
    geom_sf(data = UM_cdec) + 
    xlab(NULL)+
    ylab(NULL)+
    ggtitle("CDEC stations") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    north(sf_UM, symbol=10, location="bottomright") + 
    scalebar(UM_boundary,
             dist=10, dist_unit="km", st.size = 3,
             transform=FALSE)
```


#CDEC with R
```{r}
station = "EPW"
duration = "event" # or "E"

cdec_query(station, sensor, duration, start.date, end.date)
```


