---
title: "UpperMerced_Volumes"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(dataRetrieval)
library(ggsn)
library(sf)
library(maps)
#library(raster)
library(sp)
```

#Maps
##Get UM boundary shapefile
```{r}
getwd()

# Unzip the shapefile
#unzip("WBD_18_HU2_Shape.zip")

# Read in the shapefile with st_read()
#shapefile <- st_read("C:/Users/jessi/Desktop/SierraNevadaBeavers/SierraNevadaBeavers/Shape/WBDHU8.shp")

# Select just the Upper Merced Watershed
#UM_boundary <- st_write(shapefile[shapefile$huc8=="18040008",], "UM_boundary.shp", append = FALSE)

#remove unneeded shapefiles
#file.remove(list.files(pattern = "WBD_18_HU2_Shape*",recursive = F))

#Project to California Teal Albers (NAD83) EPSG code 3310
#UM_boundary <- st_transform(UM_boundary, 3310)

#saveRDS(UM_boundary,"UM_boundary")
UM_boundary <- readRDS("UM_boundary")

```
##UM Boundary Map
```{r}
ggplot() +
    geom_sf(data = UM_boundary) +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle("Upper Merced)") + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    north(UM_boundary, symbol=10, location="bottomright") + 
    scalebar(UM_boundary,
             dist=10, dist_unit="km", st.size = 3,
             transform=FALSE)
```


#Import Practice Shapefile

```{r}
getwd()

# Read in the shapefile with st_read()
#damCapClip <- st_read("C:/Users/jessie_moravek/Desktop/SierraNevadaBeavers/DamCapacity_Clip.shp")

#remove unneeded shapefiles
#file.remove(list.files(pattern = "DamCapacity_Clip*",recursive = F))

#Project to California Teal Albers (NAD83) EPSG code 3310
#damCapClip <- st_transform(damCapClip, 3310)

#saveRDS(damCapClip,"damCapClip.shp")
damCapClip <- readRDS("damCapClip.shp")
```
##Visualize
```{r}
ggplot() +
    geom_sf(data = damCapClip) +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle("Dam Capacity Clipped Area") + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    north(damCapClip, symbol=10, location="bottomright") + 
    scalebar(damCapClip,
             dist=.5, dist_unit="km", st.size = 3,
             transform=FALSE)
```


##Attribute table
```{r}
names(damCapClip)
damCapClip$iGeo_Slope
```
#Volume and Surface Area calculation
```{r}
damCapClip_v <- damCapClip
for (i in damCapClip_v){ #for each stream segment in the BRAT model output shapefile
  damCt <- damCapClip_v$mCC_EX_CT #number of beaver dams per segment (calculated based on dam density and segment length as the column mCC_EX_CT in the attribute table)
  slope <- damCapClip_v$iGeo_Slope #meters
  da <- damCapClip_v$iGeo_DA #m2
  width <- ifelse(da < 4.95,2.18*(da)^0.191,
                  ifelse(da >= 4.95 & da < 377,1.41*(da)^0.462,7.18*(da)^0.183)) #based on Scamardo et al. 2022
  height <- 1 #assume all beaver dams are 1m tall
  length <- height/slope #trigonometry: slope = rise/run, so run = rise/slope
  maxLength<- (damCapClip_v$iGeo_Len/damCt) #set a maximum length so that we can change volume calculations if there are so many dams that they all run together
  volume <- ifelse(length < maxLength, 0.5*(height)*(width)*(length)*(damCt),(height)*(width)*(maxLength)*(damCt))
  surfaceArea <- ifelse(length < maxLength, (width)*(length)*(damCt),(width)*(maxLength)*(damCt))
  damCapClip_v$width <- width
  damCapClip_v$volume <- volume #m3
  damCapClip_v$surfaceArea <- surfaceArea #m2
} 

saveRDS(damCapClip_v,"damCapClip_v.rds")
```


##Create 1km grid
```{r}
#Check if sf object is in metric CRS (i.e. projected)
st_is_longlat(damCapClip)
#st_is_longlat returns "False" which means that the object damCapClip is in a metric CRS

#CRS 3310 is in units of meters (see here https://epsg.io/3310)
grid_spacing <- 1000 

grid1 <- st_make_grid(damCapClip, square = T, cellsize = c(grid_spacing, grid_spacing)) %>% # the grid, covering bounding box
  st_sf()

#add an id column
grid1 <- mutate(grid1, id = 1:nrow(grid1))
```




##Volume by grid cell
```{r}
intersection <- st_intersection(grid1, damCapClip_v) %>% 
  mutate(length = st_length(.)) %>% #creates a "length" column that is the length of that particular segment in the grid cell from the "id" column
  st_drop_geometry() # complicates things in joins later on

grid1 <- grid1 %>% 
  left_join(intersection, by = "id")

vol_by_grid <- grid1 %>% 
  mutate(percent_len_in_cell = length/iGeo_Len) %>% 
  mutate(segment_vol_in_cell = percent_len_in_cell*volume) %>% 
  group_by(id) %>% summarize(total_vol_in_cell = sum(segment_vol_in_cell))


vol_by_grid$total_vol_in_cell <- as.numeric(vol_by_grid$total_vol_in_cell) #convert volume col to numeric
```
```{r}
ggplot() +
    geom_sf(data = vol_by_grid, aes(fill = total_vol_in_cell), color = NA) +
    geom_sf(data = damCapClip, col = "blue") +
    scale_fill_gradient2(low = "#132B43", high = "#078FF7") +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle( expression(paste("Volume per ", km^2))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5), 
          panel.grid.major = element_line(colour = "transparent")) +
    north(damCapClip, symbol=10, location="bottomleft") + 
    scalebar(vol_by_grid,
             dist=.5, dist_unit="km", st.size = 3, transform= F) +
    guides(fill = guide_colorbar(title = expression(paste(m^3)),
                             title.position = "bottom",
                             title.theme = element_text(size = 10,
                                                        colour = "gray70",
                                                        angle = 0)))

summary(vol_by_grid$total_vol_in_cell,na.rm=T)
summary(damCapClip$mCC_EX_CT)
```

##Surface Area by grid cell
```{r}
#use grid1 (after doing the leftjoin above for volume calculations)

sa_by_grid <- grid1 %>% 
  mutate(percent_len_in_cell = length/iGeo_Len) %>% 
  mutate(segment_sa_in_cell = percent_len_in_cell*surfaceArea) %>% 
  group_by(id) %>% summarize(total_sa_in_cell = sum(segment_sa_in_cell))

sa_by_grid
sa_by_grid$total_sa_in_cell <- as.numeric(sa_by_grid$total_sa_in_cell) #convert volume col to numeric
```

```{r}
ggplot() +
    geom_sf(data = sa_by_grid, aes(fill = total_sa_in_cell), color = NA) +
    geom_sf(data = damCapClip, col = "blue") +
    scale_fill_gradient2(low = "#43132B", high = "#F7078F") +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle( expression(paste("Surface Area per ", km^2))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5), 
          panel.grid.major = element_line(colour = "transparent")) +
    north(damCapClip, symbol=10, location="bottomleft") + 
    scalebar(vol_by_grid,
             dist=.5, dist_unit="km", st.size = 3, transform= F) +
    guides(fill = guide_colorbar(title = expression(paste(m^2)),
                             title.position = "bottom",
                             title.theme = element_text(size = 10,
                                                        colour = "gray70",
                                                        angle = 0)))

median(vol_by_grid$total_vol_in_cell,na.rm=T)
```

#Upper Merced
##Bounding Box
```{r}
#This box aligns with this LIS model coordinates
#getwd()

# Read in the shapefile with st_read()
#boundingBox <- st_read("C:/Users/jessie_moravek/Desktop/SierraNevadaBeavers/BoundingBox.shp")

#remove unneeded shapefiles
#file.remove(list.files(pattern = "BoundingBox*",recursive = F))

#Project to California Teal Albers (NAD83) EPSG code 3310
#boundingBox <- st_transform(boundingBox, 3310)

#saveRDS(boundingBox,"boundingBox")
boundingBox <- readRDS("boundingBox")
```
##Dam Capacity
```{r}
#This box aligns with this LIS model coordinates
#getwd()

# Read in the shapefile with st_read()
#damCapacity <- st_read("C:/Users/jessie_moravek/Desktop/SierraNevadaBeavers/DamCapacity.shp")

#remove unneeded shapefiles
#file.remove(list.files(pattern = "damCapacity*",recursive = F))

#Project to California Teal Albers (NAD83) EPSG code 3310
#damCapacity <- st_transform(damCapacity, 3310)

#saveRDS(damCapacity,"damCapacity")
damCapacity <- readRDS("damCapacity")
```
##Create 1km grid
```{r}
#Check if sf object is in metric CRS (i.e. projected)
st_is_longlat(boundingBox)
#st_is_longlat returns "False" which means that the object is in a metric CRS

#CRS 3310 is in units of meters (see here https://epsg.io/3310)
grid_spacing <- 1000 

grid2 <- st_make_grid(boundingBox, square = T, cellsize = c(grid_spacing, grid_spacing)) %>% # the grid, covering bounding box
  st_sf()

#add an id column
grid2 <- mutate(grid2, id = 1:nrow(grid2))
```

###Map
```{r}
ggplot() +
    geom_sf(data = grid2, fill = NA) +
    geom_sf(data = damCapacity, col = "blue") +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle( expression(paste("Upper Merced 1km Grid"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5), 
          panel.grid.major = element_line(colour = "transparent")) +
    north(damCapacity, symbol=10, location="bottomright") + 
    scalebar(damCapacity,
             dist=10, dist_unit="km", st.size = 3, transform= F) 
```
##Volume and Surface Area calculation
```{r}
damCapacity_v <- damCapacity
for (i in damCapacity_v){ #for each stream segment in the BRAT model output shapefile
  damDens <- damCapacity_v$mCC_EX_CT #number of beaver dams per segment (calculated based on dam density and segment length as the column mCC_EX_CT in the attribute table)
  slope <- damCapacity_v$iGeo_Slope #meters
  da <- damCapacity_v$iGeo_DA #m2
  width <- ifelse(da < 4.95,2.18*(da)^0.191,
                  ifelse(da >= 4.95 & da < 377,1.41*(da)^0.462,7.18*(da)^0.183)) #based on Scamardo et al. 2022
  length <- 1/slope 
  maxLength<- (damCapacity_v$iGeo_Len/damCapacity_v$mCC_EX_CT)
    volume <- ifelse(length < maxLength, 0.5*1*(width)*(length)*(damDens),1*(width)*(maxLength)*(damDens))
  surfaceArea <- ifelse(length < maxLength, (width)*(length)*(damDens),(width)*(maxLength)*(damDens))
  damCapacity_v$width <- width
  damCapacity_v$volume <- volume #m3
  damCapacity_v$surfaceArea <- surfaceArea #m2
} 

saveRDS(damCapacity_v,"damCapacity_v.rds")
```

##Volume by grid cell
```{r}
intersection2 <- st_intersection(grid2, damCapacity_v) %>% 
  mutate(length = st_length(.)) %>% #creates a "length" column that is the length of that particular segment in the grid cell from the "id" column
  st_drop_geometry() # complicates things in joins later on

grid2 <- grid2 %>% 
  left_join(intersection2, by = "id")

vol_by_grid2 <- grid2 %>% 
  mutate(percent_len_in_cell = length/iGeo_Len) %>% 
  mutate(segment_vol_in_cell = percent_len_in_cell*volume) %>% 
  group_by(id) %>% summarize(total_vol_in_cell = sum(segment_vol_in_cell))

vol_by_grid2$total_vol_in_cell <- as.numeric(vol_by_grid2$total_vol_in_cell) #convert volume col to numeric
saveRDS(vol_by_grid2,"vol_by_grid2.rds")
```
###Map
```{r}
ggplot() +
    geom_sf(data = vol_by_grid2, aes(fill = total_vol_in_cell), color = NA) +
    geom_sf(data = damCapacity, col = "blue") +
    scale_fill_gradient2(limits = c(0.01,76810.52),low = "#132B43", high = "#078FF7") +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle( expression(paste("Volume per ", km^2))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5), 
          panel.grid.major = element_line(colour = "transparent")) +
    north(damCapacity, symbol=10, location="bottomleft") + 
    scalebar(vol_by_grid2,
             dist=20, dist_unit="km", st.size = 3, transform= F) +
    guides(fill = guide_colorbar(title = expression(paste(m^3)),
                             title.position = "bottom",
                             title.theme = element_text(size = 10,
                                                        colour = "gray70",
                                                        angle = 0)))
0:150
?scale_fill_gradient2
summary(vol_by_grid2$total_vol_in_cell,na.rm=T)
```
##Surface Area by grid cell
```{r}
#use grid2 (after doing the leftjoin above for volume calculations)

sa_by_grid2 <- grid2 %>% 
  mutate(percent_len_in_cell = length/iGeo_Len) %>% 
  mutate(segment_sa_in_cell = percent_len_in_cell*surfaceArea) %>% 
  group_by(id) %>% summarize(total_sa_in_cell = sum(segment_sa_in_cell))

sa_by_grid2$total_sa_in_cell <- as.numeric(sa_by_grid2$total_sa_in_cell) #convert volume col to numeric

saveRDS(sa_by_grid2,"sa_by_grid2")
```
###Map
```{r}
ggplot() +
    geom_sf(data = sa_by_grid2, aes(fill = total_sa_in_cell), color = NA) +
    #geom_sf(data = damCapacity, col = "blue") +
    scale_fill_gradient2(limits = c(0,10000), low = "#43132B", high = "#F7078F") +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle( expression(paste("Surface Area per ", km^2))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5), 
          panel.grid.major = element_line(colour = "transparent")) +
    north(damCapacity, symbol=10, location="bottomleft") + 
    scalebar(vol_by_grid2,
             dist=20, dist_unit="km", st.size = 3, transform= F) +
    guides(fill = guide_colorbar(title = expression(paste(m^2)),
                             title.position = "bottom",
                             title.theme = element_text(size = 10,
                                                        colour = "gray70",
                                                        angle = 0)))
0:150
?scale_fill_gradient2
summary(sa_by_grid2$total_sa_in_cell,na.rm=T)
```

#Investigate outliers
```{r}
vol_by_grid2
sa_by_grid2

hist(vol_by_grid2$total_vol_in_cell)
summary(vol_by_grid2$total_vol_in_cell)

vol_by_grid2 %>% arrange(desc(total_vol_in_cell))
```
```{r}
ggplot() +
    geom_sf(data = vol_by_grid2, aes(fill = total_vol_in_cell), color = NA) +
    geom_sf(data = damCapacity, col = "blue") +
    scale_fill_gradient2(limits = c(10000,52628.57442), low = "#43132B", high = "#F7078F") +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle( expression(paste("Volume per ", km^2))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5), 
          panel.grid.major = element_line(colour = "transparent")) +
    north(damCapacity, symbol=10, location="bottomleft") + 
    scalebar(vol_by_grid2,
             dist=20, dist_unit="km", st.size = 3, transform= F) +
    guides(fill = guide_colorbar(title = expression(paste(m^3)),
                             title.position = "bottom",
                             title.theme = element_text(size = 10,
                                                        colour = "gray70",
                                                        angle = 0)))
```
#Re-do Dam Capacity
##Note that we weren't using the correct max DA
##Bounding Box
```{r}
#This box aligns with this LIS model coordinates
#getwd()

# Read in the shapefile with st_read()
#boundingBox2 <- st_read("C:/Users/jessie_moravek/Desktop/SierraNevadaBeavers/BoundingBox2.shp")

#remove unneeded shapefiles
#file.remove(list.files(pattern = "BoundingBox*",recursive = F))

#Project to California Teal Albers (NAD83) EPSG code 3310
boundingBox2 <- readRDS("boundingBox2")

boundingBox2 <- st_transform(boundingBox2, 3310)

saveRDS(boundingBox2,"boundingBox2")
```
##Dam Capacity
```{r}
#This box aligns with this LIS model coordinates
#getwd()

# Read in the shapefile with st_read()
#damCapacity2 <- st_read("C:/Users/jessie_moravek/Desktop/SierraNevadaBeavers/DamCapacity2.shp")

#remove unneeded shapefiles
#file.remove(list.files(pattern = "damCapacity2*",recursive = F))

#Project to California Teal Albers (NAD83) EPSG code 3310
damCapacity2 <- readRDS("damCapacity2")

damCapacity2 <- st_transform(damCapacity2, 3310)
saveRDS(damCapacity2,"damCapacity2")
```

##Create 1km grid
```{r}
#Check if sf object is in metric CRS (i.e. projected)
st_is_longlat(boundingBox2)
#st_is_longlat returns "False" which means that the object is in a metric CRS

#CRS 3310 is in units of meters (see here https://epsg.io/3310)
grid_spacing <- 1000 

grid3 <- st_make_grid(boundingBox2, square = T, cellsize = c(grid_spacing, grid_spacing)) %>% # the grid, covering bounding box
  st_sf()

#add an id column
grid3 <- mutate(grid3, id = 1:nrow(grid3))
dim(grid3)
```

###Map
```{r}
ggplot() +
    geom_sf(data = grid3, fill = NA) +
    geom_sf(data = damCapacity2, col = "blue") +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle( expression(paste("Upper Merced 1km Grid"))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5), 
          panel.grid.major = element_line(colour = "transparent")) +
    north(damCapacity2, symbol=10, location="bottomright") + 
    scalebar(damCapacity2,
             dist=10, dist_unit="km", st.size = 3, transform= F) 
```

##Volume and Surface Area calculation
```{r}
damCapacity2_v <- damCapacity2
for (i in damCapacity2_v){ #for each stream segment in the BRAT model output shapefile
  damCt <- damCapacity2_v$mCC_EX_CT #number of beaver dams per segment (calculated based on dam density and segment length as the column mCC_EX_CT in the attribute table)
  slope <- damCapacity2_v$iGeo_Slope #meters
  da <- damCapacity2_v$iGeo_DA #m2
  width <- ifelse(da < 4.95,2.18*(da)^0.191,
                  ifelse(da >= 4.95 & da < 377,1.41*(da)^0.462,7.18*(da)^0.183)) #based on Scamardo et al. 2022
  height <- 1 #assume all beaver dams are 1m tall
  length <- height/slope #trigonometry: slope = rise/run, so run = rise/slope
  maxLength<- (damCapacity2_v$iGeo_Len/damCt) #set a maximum length so that we can change volume calculations if there are so many dams that they all run together
  volume <- ifelse(length < maxLength, 0.5*(height)*(width)*(length)*(damCt),(height)*(width)*(maxLength)*(damCt))
  surfaceArea <- ifelse(length < maxLength, (width)*(length)*(damCt),(width)*(maxLength)*(damCt))
  damCapacity2_v$width <- width
  damCapacity2_v$volume <- volume #m3
  damCapacity2_v$surfaceArea <- surfaceArea #m2
} 

saveRDS(damCapacity2_v,"damCapacity2_v.rds")
```

##Volume by grid cell
```{r}
intersection3 <- st_intersection(grid3, damCapacity2_v) %>% 
  mutate(length = st_length(.)) %>% #creates a "length" column that is the length of that particular segment in the grid cell from the "id" column
  st_drop_geometry() # complicates things in joins later on

grid3 <- grid3 %>% 
  left_join(intersection3, by = "id")

vol_by_grid3 <- grid3 %>% 
  mutate(percent_len_in_cell = length/iGeo_Len) %>% 
  mutate(segment_vol_in_cell = percent_len_in_cell*volume) %>% 
  group_by(id) %>% summarize(total_vol_in_cell = sum(segment_vol_in_cell))

vol_by_grid3$total_vol_in_cell <- as.numeric(vol_by_grid3$total_vol_in_cell) #convert volume col to numeric
saveRDS(vol_by_grid3,"vol_by_grid3.rds")
vol_by_grid3 <- readRDS("vol_by_grid3.rds")
sum(vol_by_grid3$total_vol_in_cell, na.rm=T)

```
###Map
```{r}
ggplot() +
    geom_sf(data = vol_by_grid3, aes(fill = total_vol_in_cell), color = NA) +
    #geom_sf(data = damCapacity2, col = "blue") +
    scale_fill_gradient2(low = "gray70", high = "#078FF7") +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle( expression(paste("Volume per ", km^2))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5), 
          panel.grid.major = element_line(colour = "transparent")) +
    north(damCapacity2, symbol=10, location="bottomleft") + 
    scalebar(vol_by_grid3,
             dist=20, dist_unit="km", st.size = 3, transform= F) +
    guides(fill = guide_colorbar(title = expression(paste(m^3)),
                             title.position = "bottom",
                             title.theme = element_text(size = 10,
                                                        colour = "gray70",
                                                        angle = 0)))

?scale_fill_gradient2
summary(vol_by_grid3$total_vol_in_cell,na.rm=T)
```






##Surface Area by grid cell
```{r}
#use grid3 (after doing the leftjoin above for volume calculations)

sa_by_grid3 <- grid3 %>% 
  mutate(percent_len_in_cell = length/iGeo_Len) %>% 
  mutate(segment_sa_in_cell = percent_len_in_cell*surfaceArea) %>% 
  group_by(id) %>% summarize(total_sa_in_cell = sum(segment_sa_in_cell))

sa_by_grid3$total_sa_in_cell <- as.numeric(sa_by_grid3$total_sa_in_cell) #convert volume col to numeric

saveRDS(sa_by_grid3,"sa_by_grid3.rds")
sa_by_grid3 <- readRDS('sa_by_grid3.rds')

```
###Map
```{r}
ggplot() +
    geom_sf(data = sa_by_grid3, aes(fill = total_sa_in_cell), color = NA) +
    #geom_sf(data = damCapacity2, col = "blue") +
    scale_fill_gradient2(low = "gray70", high = "#F7078F") +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle( expression(paste("Surface Area per ", km^2))) + 
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5), 
          panel.grid.major = element_line(colour = "transparent")) +
    north(damCapacity2, symbol=10, location="bottomleft") + 
    scalebar(vol_by_grid3,
             dist=20, dist_unit="km", st.size = 3, transform= F) +
    guides(fill = guide_colorbar(title = expression(paste(m^2)),
                             title.position = "bottom",
                             title.theme = element_text(size = 10,
                                                        colour = "gray70",
                                                        angle = 0)))

?scale_fill_gradient2
summary(sa_by_grid3$total_sa_in_cell,na.rm=T)
```
##Reproject for rasterization 
```{r}
vol_by_grid3_wgs <- st_transform(vol_by_grid3, 4326)
sa_by_grid3_wgs <- st_transform(sa_by_grid3, 4087)

saveRDS(vol_by_grid3_wgs, "vol_by_grid3_wgs.rds")
saveRDS(sa_by_grid3_wgs, "sa_by_grid3_wgs.rds")
```










#FINAL GRID IN CORRECT NASA CRS
```{r}
library(sf)
library(sp)
library(raster)
library(lwgeom)
```
##NASA projection
```{r}
wkt <- 'GEOGCS["WGS 84",DATUM["WGS_1984",SPHEROID["WGS 84",6378137,298.257223563,AUTHORITY["EPSG","7030"]],AUTHORITY["EPSG","6326"]],PRIMEM["Greenwich",0,AUTHORITY["EPSG","8901"]],UNIT["degree",0.0174532925199433,AUTHORITY["EPSG","9122"]],AXIS["Latitude",NORTH],AXIS["Longitude",EAST],AUTHORITY["EPSG","4326"]]'
```


##Dam Capacity
```{r}
#This box aligns with this LIS model coordinates
#getwd()

# Read in the shapefile with st_read()
#damCapacity2 <- st_read("C:/Users/jessie_moravek/Desktop/SierraNevadaBeavers/DamCapacity2.shp")

#remove unneeded shapefiles
#file.remove(list.files(pattern = "damCapacity2*",recursive = F))

#Project to California Teal Albers (NAD83) EPSG code 3310
damCapacity2 <- readRDS("damCapacity2")

damCapacity2 <- st_transform(damCapacity2, wkt)
saveRDS(damCapacity2,"damCapacity2")
```

##Making a grid
```{r}
xmin <- -120.985
xmax <- -119.255
ymin <- 37.325
ymax <- 37.925
extent <- extent(xmin, xmax, ymin, ymax)

study_area2 <- st_bbox(extent, crs = st_crs(wkt))

study_area3 = st_bbox(st_transform(st_as_sfc(study_area2), crs = wkt))
#cellsize = (c(diff(study_area3[c(1, 3)]), diff(study_area3[c(2, 4)])))/n, 

study_area4 = st_make_grid(study_area3, n = c(174,63))

study_area5 = st_as_sf(study_area4)
class(study_area5)

#add an id column
study_area5$id=1:length(study_area4)
```
##Volume and Surface Area calculation
```{r}
damCapacity2_v <- damCapacity2
for (i in damCapacity2_v){ #for each stream segment in the BRAT model output shapefile
  damCt <- damCapacity2_v$mCC_EX_CT #number of beaver dams per segment (calculated based on dam density and segment length as the column mCC_EX_CT in the attribute table)
  slope <- damCapacity2_v$iGeo_Slope #meters
  da <- damCapacity2_v$iGeo_DA #m2
  width <- ifelse(da < 4.95,2.18*(da)^0.191,
                  ifelse(da >= 4.95 & da < 377,1.41*(da)^0.462,7.18*(da)^0.183)) #based on Scamardo et al. 2022
  height <- 1 #assume all beaver dams are 1m tall
  length <- height/slope #trigonometry: slope = rise/run, so run = rise/slope
  maxLength<- (damCapacity2_v$iGeo_Len/damCt) #set a maximum length so that we can change volume calculations if there are so many dams that they all run together
  volume <- ifelse(length < maxLength, 0.5*(height)*(width)*(length)*(damCt),(height)*(width)*(maxLength)*(damCt))
  surfaceArea <- ifelse(length < maxLength, (width)*(length)*(damCt),(width)*(maxLength)*(damCt))
  damCapacity2_v$width <- width
  damCapacity2_v$volume <- volume #m3
  damCapacity2_v$surfaceArea <- surfaceArea #m2
} 


saveRDS(damCapacity2_v,"damCapacity2_v.rds")
```

##Volume by grid cell
```{r}
sf_use_s2(FALSE)
intersection5 <- st_intersection(study_area5, damCapacity2_v) %>% 
  mutate(length = st_length(.)) %>% #creates a "length" column that is the length of that particular segment in the grid cell from the "id" column
  st_drop_geometry() # complicates things in joins later on

study_area5 <- study_area5 %>% 
  left_join(intersection5, by = "id")

vol_by_grid5 <- study_area5 %>% 
  mutate(percent_len_in_cell = length/iGeo_Len) %>% 
  mutate(segment_vol_in_cell = percent_len_in_cell*volume) %>% 
  group_by(id) %>% summarize(total_vol_in_cell = sum(segment_vol_in_cell))

vol_by_grid5$total_vol_in_cell <- as.numeric(vol_by_grid5$total_vol_in_cell) #convert volume col to numeric

saveRDS(vol_by_grid5,"vol_by_grid5.rds")
vol_by_grid5 <- readRDS("vol_by_grid5.rds")

```

###Map
```{r}

ggplot() +
    geom_sf(data = vol_by_grid5, aes(fill = total_vol_in_cell), color = NA) +
    #geom_sf(data = damCapacity2, col = "blue") +
    scale_fill_gradient2(low = "gray70", high = "#078FF7") +
    xlab(NULL)+
    ylab(NULL)+
    ggtitle( expression(paste("Volume per ", km^2))) + 
    theme_minimal() 
```
```{r}
vol_by_grid5
```


##make NetCDF file
```{r}
st_write(vol_by_grid5, "vol_by_grid5.nc")
```


##Writing the code in R instead of Python

```{r}
library(ncdf4)
library(terra)
```

###Practice code
```{r}
# Create a small 2x3x4 array
height_array <- array(cumsum(abs(rnorm(24))), dim = c(5, 5, 10))

# Print the array
print(height_array)
height_1 <- height_array[1,1,]

#multiply by arrea

area_matrix <- matrix(cumsum(abs(rnorm(24))), 5, 5)

area_1 <- area_matrix[1,1]

volume_1 <- height_1 * area_1
plot(height_1, volume_1)

model_1<-lm(volume_1 ~ height_1)

m <- coef(model_1)[2]
b <- coef(model_1)[1]

damvol <- 3

floodheight <- (damvol - b )/m

for (i in 1:nrow(area_matrix)){
  for (j in 1:ncol(area_matrix)){
    area = area_matrix[i,j]
    height = height_array[i,j,]
    volume = area * height
    model = lm(volume ~ height)
    m = coef(model)[2]
    b = coef(model)[1]
    damvol = 2
    floodheight = (damvol-b)/m
    print(floodheight)
 }}   
    

```


###Data we need
```{r}
#volumes
rvol <- vol_by_grid5
?rast
#make it a raster
rvol <- rast(ncols = 174, nrows = 63, xmin = -120.985, xmax = -119.255, ymin = 37.325, ymax = 37.925)
rvol
values(rvol) <- vol_by_grid5$total_vol_in_cell 
plot(rvol,1, colNA="grey70") #hmm upside down
rvol  <- flip(rvol, direction="vertical")
plot(rvol,1, colNA="grey70")

```


```{r}
#lis inputs
inputs <- nc_open("lis_input.nc")
print(inputs)

attributes(inputs$var)
attributes(inputs$dim)

# Get latitude and longitude with the ncvar_get function and store each into their own object:
lat <- ncvar_get(inputs, "lat")
nlat <- dim(lat) #to check it matches the metadata: 23
lon <- ncvar_get(inputs, "lon")
nlon <- dim(lon) #to check, should be 24
# Check your lat lon dimensions match the information in the metadata we explored before:
print(c(nlon, nlat))
latlon <- ncvar_get(inputs, "lat")

#now pull an array
HYMAP_grid_area <- ncvar_get(inputs, "HYMAP_grid_area")

```

```{r}
for (i in 1:nrow(inputs)){
  for (j in 1:ncol(inputs)){
    area = area_matrix[i,j]
    height = height_array[i,j,]
    volume = area * height
    model = lm(volume ~ height)
    m = coef(model)[2]
    b = coef(model)[1]
    damvol = 2
    floodheight = (damvol-b)/m
    print(floodheight)
 }}  
```




